### 2014-04-22 paper/paper.tex
Initial commit - paper and code files

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09

\title{Formatting Instructions for NIPS 2014}

\author{
Stephanie deWet \\
Department of Computer Sciences\\
University of Wisconsin - Madison \\
Madison, WI 53705 \\
\texttt{sdewet@cs.wisc.edu} \\
\And
Saswati De \\
Department of Computer Sciences\\
University of Wisconsin - Madison \\
Madison, WI 53705 \\
\texttt{@cs.wisc.edu} \\

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}
Financial data is fun.

\section{Initial Data Processing}
Our data set consists of information collected on a single stock, including instantaneous stock prices and instantaneous values of 26 parameters ($P$) known to have some affect on the price.
These values are taken at many different time stamps.
Our goal is to predict the stock price in the near future, given current values of the parameters.

First, we need to create labeled data to use in training our models.
We use two strategies to do this.

\subsection{Strategy 1}
We sort the data set in increasing time.
For each line, we define $\delta$, the future change in stock price, to be the current stock price subtracted from the next stock price in the future.
We will attempt to build a model which calculates $\delta_t$ when given $P_t$, the values of the 26 parameters at time $t$.

% Note to self: this is done in strategy

\subsubsection*{Acknowledgments}
Thanks to Jerry Zhu for an excellent class, and to Rasmussen and Willams for their excellent book.

\subsubsection*{References}

\small{
[1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
and T.K. Leen (eds.), {\it Advances in Neural Information Processing
Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.

[2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
Realistic Neural Models with the GEneral NEural SImulation System.}
New York: TELOS/Springer-Verlag.

[3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
and recall at excitatory recurrent synapses and cholinergic modulation
in rat hippocampal region CA3. {\it Journal of Neuroscience}
{\bf 15}(7):5249-5262.
}

\end{document}

### 2014-04-22 code/readData.m

@@ -1,0 +1,17 @@
+function data = readData()
+% Reads in data from the file and stores it as a matrix.
+
+fileName = 'prod_data_v.txt';
+path = '../workdir/';
+
+fileID = fopen([path fileName],'r');
+formatSpec = '%d';
+nCols = 33;
+for i = 2:nCols
+    formatSpec = [formatSpec ' %f'];
+end
+
+data = fscanf(fileID, formatSpec, [nCols Inf]);
+data = data';
+
+fclose(fileID);

### 2014-04-22 code/readRegFile.m

@@ -1,0 +1,17 @@
+function reg_data = readRegFile()
+% Reads in data from the file and stores it as a matrix.
+
+fileName = 'outvals';
+path = '../workdir/sample_model_build_runs/lm/';
+
+fileID = fopen([path fileName],'r');
+formatSpec = '%d';
+nCols = 33-6+1; 
+for i = 2:nCols
+    formatSpec = [formatSpec ' %f'];
+end
+
+reg_data = fscanf(fileID, formatSpec, [nCols Inf]);
+reg_data = reg_data';
+
+fclose(fileID);

### 2014-04-22 setup.sh

@@ -1,0 +1,1 @@
+unzip workdir.zip

### 2014-04-23 code/readProcessedData.m
Raw data processing and setup scripts
This adds a script and matlab code for reading the processed data.

This also adds a readme file, and a setup script which sets up the
directory and processes the data.

@@ -1,0 +1,20 @@
+function [y, X] = readProcessedFile()
+% Reads in data from the file and stores it as a matrix.
+addpath '../data/';
+
+fileName = 'processed_data.txt';
+
+fileID = fopen(fileName,'r');
+formatSpec = '%f';
+nCols = 33-6+1;
+for i = 2:nCols
+    formatSpec = [formatSpec ' %f'];
+end
+
+data = fscanf(fileID, formatSpec, [nCols Inf]);
+data = data';
+
+fclose(fileID);
+
+y = data(:,1);
+X = data(:,2:nCols);

### 2014-04-23 code/readRegFile.m

@@ -1,17 +1,0 @@
-function reg_data = readRegFile()
-% Reads in data from the file and stores it as a matrix.
-
-fileName = 'outvals';
-path = '../workdir/sample_model_build_runs/lm/';
-
-fileID = fopen([path fileName],'r');
-formatSpec = '%d';
-nCols = 33-6+1; 
-for i = 2:nCols
-    formatSpec = [formatSpec ' %f'];
-end
-
-reg_data = fscanf(fileID, formatSpec, [nCols Inf]);
-reg_data = reg_data';
-
-fclose(fileID);

### 2014-04-23 setup.sh

@@ -1,1 +1,0 @@
-unzip workdir.zip

### 2014-04-23 paper/paper.tex
Explain raw data processing in paper

@@
 
-First, we need to create labeled data to use in training our models.
-We use two strategies to do this.
-
-\subsection{Strategy 1}
+First, we create labeled data to use in training our models using the following process:
 We sort the data set in increasing time.
-For each line, we define $\delta$, the future change in stock price, to be the current stock price subtracted from the next stock price in the future.
+For each line, we define $\delta$, the difference in stock price after 2 minutes, to be the current stock price subtracted from the stock price that is associated with the datapoint that is closest to 2 minutes in the future.
+(Note that our datapoints are roughly 50 milliseconds apart, so we can expect that the variation in distance between sets of data points will be much less than modeling error.)
 We will attempt to build a model which calculates $\delta_t$ when given $P_t$, the values of the 26 parameters at time $t$.

### 2014-04-23 paper/paper.tex
minor bugfixes

@@
 \section{Initial Data Processing}
-Our data set consists of information collected on a single stock, including instantaneous stock prices and instantaneous values of 26 parameters ($P$) known to have some affect on the price.
+Our data set consists of information collected on a single stock, including instantaneous stock prices and instantaneous values of 27 parameters ($P$) known to have some affect on the price.
 These values are taken at many different time stamps.
@@
 (Note that our datapoints are roughly 50 milliseconds apart, so we can expect that the variation in distance between sets of data points will be much less than modeling error.)
-We will attempt to build a model which calculates $\delta_t$ when given $P_t$, the values of the 26 parameters at time $t$.
+We will attempt to build a model which calculates $\delta_t$ when given $P_t$, the values of the 27 parameters at time $t$.
 
@@
 \small{
-[1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
-for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
-and T.K. Leen (eds.), {\it Advances in Neural Information Processing
-Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.
-
-[2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
-Realistic Neural Models with the GEneral NEural SImulation System.}
-New York: TELOS/Springer-Verlag.
-
-[3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
-and recall at excitatory recurrent synapses and cholinergic modulation
-in rat hippocampal region CA3. {\it Journal of Neuroscience}
-{\bf 15}(7):5249-5262.
 }

### 2014-04-23 code/readProcessedData.m

@@ -1,4 +1,4 @@
-function [y, X] = readProcessedFile()
+function [y, X] = readProcessedData()
 % Reads in data from the file and stores it as a matrix.
 addpath '../data/';
 

### 2014-04-30 code/Gaussian_Processes_Regression.m
More code...

@@ -42,6 +42,7 @@ Design_X = Scaled_X'; Design_y = Scaled_y';
 X_Window = Design_X(:,1:windowSize); y_Window = Design_y(1:windowSize);
 ctr = windowSize;
 
+Predicted_Data = [];
 while (ctr < size(Design_X,2))
     
     %Select the hyper-parameters
@@ -78,14 +79,37 @@ while (ctr < size(Design_X,2))
 
     %Predict the change in stock price for the next time instant
     
+    k_star = ones(num_examples,1);
+    x_star = Design_X(:,(ctr+1));
+    for i=1:num_examples  
+        switch cov_type
+            case 1
+                k_star(i) = Squared_Exponential(X_Window(:,i), x_star, sigma_f, l);
+            case 2
+                k_star(i) = Ornstein_Uhlenbeck(X_Window(:,i), x_star, sigma_f, l);
+            case 3
+                k_star(i) = Matern(X_Window(:,i), x_star, sigma_f, l, (5/2));
+            case 4
+                k_star(i) = Matern(X_Window(:,i), x_star, sigma_f, l, (3/2));
+            otherwise
+                error('Wrong covariance function type. Please try again!');                    
+        end
+    end
+    L = cholesky(K);
+    alpha = L'\(L\y_Window);
+    predicted_y = k_star' * alpha;      % Our predicted change in price is the mean of the gaussian posterior
+    Predicted_Data = [Predicted_Data; [predicted_y, x_star']];
+    fprintf('Finished prediction of Data Point no. %d\n', (ctr+1));
     
     %
     ctr = ctr+1;
     X_Window = [X_Window, Design_X(:,ctr)];
     X_Window(:,1)=[];
+    y_Window = [y_Window, Design_y(ctr)];
+    y_Window(1) = [];
 
-end
-end
+end %end of while
+end %end of function
 
 function k_elem = Squared_Exponential(x_p, x_q, sigma_f, l)
 l = l.^(-2); M = diag(l);

### 2014-04-30 code/Gaussian_Processes_Regression.m
Fixed few issues

@@ -95,8 +95,8 @@ while (ctr < size(Design_X,2))
                 error('Wrong covariance function type. Please try again!');                    
         end
     end
-    L = cholesky(K);
-    alpha = L'\(L\y_Window);
+    L = chol(K, 'lower');
+    alpha = L'\(L\y_Window');
     predicted_y = k_star' * alpha;      % Our predicted change in price is the mean of the gaussian posterior
     Predicted_Data = [Predicted_Data; [predicted_y, x_star']];
     fprintf('Finished prediction of Data Point no. %d\n', (ctr+1));

### 2014-04-30 code/covar_hyperparams/K_mat.m
Calculate covariance parameters - first working version
Note: the gradient has a local discontinuity at r=0.
Would it be better to not provide the gradient?

Note 2: I also tried trust-region-reflective.  It was very slow (200+
steps, and hadn't got anywhere near the expected solution when I gave
up).

@@ -1,0 +1,15 @@
+function out = K_mat(X, theta)
+% Matern covariance function matrix (objective function)
+%    X - data matrix
+%    theta - the struct of coordinates that can be messed with.
+%       Contains ll, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
+
+[d, N] = size(X);
+out = zeros(N);
+for i = 1:N
+	out(i,i) =  k_func(X(:,i), X(:,i), theta, true);
+	for j = i+1:N
+		out(i,j) = k_func(X(:,i), X(:,j), theta, false);
+		out(j,i) = out(i,j);
+	end
+end

### 2014-04-30 code/covar_hyperparams/dk_func_ll.m

@@ -1,0 +1,18 @@
+function out = dk_func_ll(x1, x2, theta, j)
+% Derivative of Matern covariance function with respect to l[j] hyperparameter
+%    x1 - first coordinate
+%    x2 - second coordinate
+%    theta - the array of coordinates that can be changed.
+%       elements 1:d :  l, a vector of scaling params
+%       element d+1: var_f (= sigma_f^2)
+%       element d+2: var_n (= sigma_n^2)
+%    j - hyperparameter
+
+D = numel(x1);
+
+d = x1 - x2;
+u = sqrt(d' * diag(1 ./ (theta(1:D) .^2)) * d);
+
+% d/dl sigma_f * exp(-u) = sigma_f * exp(u) * (-1/(2 * sqrt(sum(d^2/ll^2)))) * (-2 d_j^2 / l_j^3)
+
+out = theta(D+1) * (exp(-u) / u) * (d(j)^2 / theta(j)^3);

### 2014-04-30 code/covar_hyperparams/dk_func_sigma_f.m

@@ -1,0 +1,13 @@
+function out = dk_func_f(x1, x2, theta)
+% Derivative of Matern covariance function with respect to sigma_f hyperparameter
+%    x1 - first coordinate
+%    x2 - second coordinate
+%    theta - the array of coordinates that can be changed.
+%       elements 1:d :  l, a vector of scaling params
+%       element d+1: var_f (= sigma_f^2)
+%       element d+2: var_n (= sigma_n^2)
+
+D = numel(x1);
+d = x1 - x2;
+sigma_f = sqrt(theta(D+1));
+out = 2 * sigma_f * exp( - sqrt( d' * diag(1 ./ (theta(1:D) .^2)) * d ));

### 2014-04-30 code/covar_hyperparams/dk_func_sigma_n.m

@@ -1,13 +1,13 @@
 function out = dk_func_n(theta, on_diagonal)
 % Derivative of Matern covariance function with respect to sigma_f hyperparameter
-%    theta - the struct of coordinates that can be messed with.
-%       Contains (l, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
+%    theta - the array of coordinates that can be changed.
+%       elements 1:d :  l, a vector of scaling params
+%       element d+1: var_f (= sigma_f^2)
+%       element d+2: var_n (= sigma_n^2)
 %    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
 
-% Note: sigma_f^2 = var_f
-
 if (on_diagonal == true)
-	sigma_n = sqrt(theta.var_n);
+	sigma_n = sqrt(theta(end));
 	out = 2 * sigma_n;
 else
 	out = 0;

### 2014-04-30 code/covar_hyperparams/k_func.m

@@ -2,18 +2,21 @@ function k = k_func(x1, x2, theta, on_diagonal)
 % Matern covariance function
 %    x1 - first coordinate
 %    x2 - second coordinate
-%    theta - the struct of coordinates that can be messed with.
-%       Contains (l, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
+%    theta - the array of coordinates that can be changed.
+%       elements 1:d :  l, a vector of scaling params
+%       element d+1: var_f (= sigma_f^2)
+%       element d+2: var_n (= sigma_n^2)
 %    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
 
-
 % For now, assume nu = 1/2
+
+D = numel(x1);  % dimensions
 
 % k = theta.var_f * exp(-norm(x1 - x2)/ theta.l) + theta.var_n * [1 if on diagonal; 0 otherwise]
 if (on_diagonal == false)
 	d = x1 - x2;
-	k = theta.var_f * exp( - sqrt( d' * diag(1 ./ (theta.ll .^2)) * d ));
+	k = theta(D+1) * exp( - sqrt( d' * diag(1 ./ (theta(1:D) .^2)) * d ));
 else
 	% exp(0) = 1, so
-	k = theta.var_f + theta.var_n;
+	k = theta(D+1) + theta(D+2);
 end

### 2014-04-30 code/covar_hyperparams/k_grad.m

@@ -1,0 +1,31 @@
+function grad = k_grad(X, y, theta, K_inv)
+
+alpha = K_inv * y;
+term1 = (alpha * alpha' - K_inv);
+
+grad = zeros(numel(theta), 1);
+[d,N] = size(X);
+
+for h = 1:numel(theta)
+	dK = zeros(N,N);
+	for i = 1:N
+		for j = i:N
+			if h <= d
+                if (i == j)
+                    % So, add a tiny bit of noise to X_i.
+                    % Not sure that this is legit, really.
+                    x_temp = X(:,j)+1e-6.*randn(d,1);
+                    dK(j,i) = dk_func_ll(X(:,i), x_temp, theta, h);
+                else
+                    dK(j,i) = dk_func_ll(X(:,i), X(:,j), theta, h);
+                end
+			elseif h == d + 1
+				dK(j,i) = dk_func_sigma_f(X(:,i), X(:,j), theta);
+			elseif h == d + 2
+				dK(j,i) = dk_func_sigma_n(theta, (i==j) );
+			end
+			dK(i, j) = dK(j, i);
+		end
+	end
+	grad(h) = 0.5 * trace(term1*dK);  % Derivative of marginal likelihood 
+end

### 2014-04-30 code/covar_hyperparams/k_obj.m

@@ -1,0 +1,14 @@
+function [obj, grad] = k_obj( X, y, theta)
+% Negative Marginal likelihood (objective function to be MINimized)
+
+K = K_mat(X, theta);
+Kinv = pinv(K);
+
+[d, N] = size(X);
+obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
+
+grad = k_grad(X, y, theta, Kinv);
+
+% Use negative of obj and grad, since we have to minimize
+obj = - obj;
+grad = -1 .* grad;

### 2014-04-30 code/dk_func_ll.m

@@ -1,14 +1,0 @@
-function out = dk_func_ll(x1, x2, theta, j)
-% Derivative of Matern covariance function with respect to l[j] hyperparameter
-%    x1 - first coordinate
-%    x2 - second coordinate
-%    theta - the struct of coordinates that can be messed with.
-%       Contains (l, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
-%    j - hyperparameter
-
-d = x1 - x2;
-u = sqrt(d' * diag(1 ./ (theta.ll .^2)) * d);
-
-% d/dl exp(-u) = exp(u) * (-1/(2 * sqrt(sum(d^2/l^2)))) * (-2 d_j^2 / l_j^3)
-
-out = theta.var_f * exp(-u) / u * d(j)^2 / theta.ll^3;

### 2014-04-30 code/dk_func_sigma_f.m

@@ -1,11 +1,0 @@
-function out = dk_func_f(x1, x2, theta)
-% Derivative of Matern covariance function with respect to sigma_f hyperparameter
-%    x1 - first coordinate
-%    x2 - second coordinate
-%    theta - the struct of coordinates that can be messed with.
-%       Contains (l, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
-
-% Note: sigma_f^2 = var_f
-d = x1 - x2;
-sigma_f = sqrt(theta.var_f);
-out = 2 * sigma_f * exp( - sqrt( d' * diag(1 ./ (theta.ll .^2)) * d ));

### 2014-04-30 code/optimize_covar.m

@@ -1,0 +1,23 @@
+function [theta_f, fval] = optimize_covar(X,y)
+
+addpath covar_hyperparams/.
+
+options = optimoptions(@fmincon,...
+'GradObj','on', ...
+'Display','iter', ...
+'Algorithm','interior-point');  % or trust-region-reflective
+
+n_theta = 29;
+theta_0 = ones(n_theta, 1);
+
+Xt = X';  % Each column of X is a 27-param table entries
+%y is a column vector
+
+objective = @(theta) k_obj(Xt,y,theta);
+
+% Lower bound
+lb = [-Inf .* ones(n_theta-2,1); 0; 0] ;
+
+[theta_f,fval] = fmincon(objective,theta_0,...
+[],[],[],[],lb,[],[],options);
+

### 2014-04-30 code/covar_hyperparams/k_obj.m
Covariance param optimization - refinement
This is faster and better without the gradient!

Also, fix lower bounds - all hyperparams should be >= 0.

@@ -1,4 +1,4 @@
-function [obj, grad] = k_obj( X, y, theta)
+function obj = k_obj( X, y, theta)
 % Negative Marginal likelihood (objective function to be MINimized)
 
 K = K_mat(X, theta);
@@ -7,8 +7,8 @@ Kinv = pinv(K);
 [d, N] = size(X);
 obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
 
-grad = k_grad(X, y, theta, Kinv);
+%grad = k_grad(X, y, theta, Kinv);
 
 % Use negative of obj and grad, since we have to minimize
 obj = - obj;
-grad = -1 .* grad;
+%grad = -1 .* grad;

### 2014-04-30 code/optimize_covar.m

@@ -1,23 +1,22 @@
-function [theta_f, fval] = optimize_covar(X,y)
+function [theta, fval] = optimize_covar(X,y)
 
 addpath covar_hyperparams/.
 
 options = optimoptions(@fmincon,...
-'GradObj','on', ...
 'Display','iter', ...
 'Algorithm','interior-point');  % or trust-region-reflective
 
 n_theta = 29;
 theta_0 = ones(n_theta, 1);
 
-Xt = X';  % Each column of X is a 27-param table entries
+Xt = X'; % Each column of X is a 27-param table entries
 %y is a column vector
 
 objective = @(theta) k_obj(Xt,y,theta);
 
-% Lower bound
-lb = [-Inf .* ones(n_theta-2,1); 0; 0] ;
+% Lower bound - bound all to be >= 0
+lb = zeros(n_theta,1);
 
-[theta_f,fval] = fmincon(objective,theta_0,...
+[theta, fval] = fmincon(objective,theta_0,...
 [],[],[],[],lb,[],[],options);
 

### 2014-04-30 code/covar_hyperparams/K_mat.m
Clean up hyperparam optimization code significantly and prepare to add other functions

@@ -1,15 +1,0 @@
-function out = K_mat(X, theta)
-% Matern covariance function matrix (objective function)
-%    X - data matrix
-%    theta - the struct of coordinates that can be messed with.
-%       Contains ll, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
-
-[d, N] = size(X);
-out = zeros(N);
-for i = 1:N
-	out(i,i) =  k_func(X(:,i), X(:,i), theta, true);
-	for j = i+1:N
-		out(i,j) = k_func(X(:,i), X(:,j), theta, false);
-		out(j,i) = out(i,j);
-	end
-end

### 2014-04-30 code/covar_hyperparams/dk_func_ll.m

@@ -1,18 +1,0 @@
-function out = dk_func_ll(x1, x2, theta, j)
-% Derivative of Matern covariance function with respect to l[j] hyperparameter
-%    x1 - first coordinate
-%    x2 - second coordinate
-%    theta - the array of coordinates that can be changed.
-%       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
-%    j - hyperparameter
-
-D = numel(x1);
-
-d = x1 - x2;
-u = sqrt(d' * diag(1 ./ (theta(1:D) .^2)) * d);
-
-% d/dl sigma_f * exp(-u) = sigma_f * exp(u) * (-1/(2 * sqrt(sum(d^2/ll^2)))) * (-2 d_j^2 / l_j^3)
-
-out = theta(D+1) * (exp(-u) / u) * (d(j)^2 / theta(j)^3);

### 2014-04-30 code/covar_hyperparams/dk_func_sigma_f.m

@@ -1,13 +1,0 @@
-function out = dk_func_f(x1, x2, theta)
-% Derivative of Matern covariance function with respect to sigma_f hyperparameter
-%    x1 - first coordinate
-%    x2 - second coordinate
-%    theta - the array of coordinates that can be changed.
-%       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
-
-D = numel(x1);
-d = x1 - x2;
-sigma_f = sqrt(theta(D+1));
-out = 2 * sigma_f * exp( - sqrt( d' * diag(1 ./ (theta(1:D) .^2)) * d ));

### 2014-04-30 code/covar_hyperparams/dk_func_sigma_n.m

@@ -1,14 +1,0 @@
-function out = dk_func_n(theta, on_diagonal)
-% Derivative of Matern covariance function with respect to sigma_f hyperparameter
-%    theta - the array of coordinates that can be changed.
-%       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
-%    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
-
-if (on_diagonal == true)
-	sigma_n = sqrt(theta(end));
-	out = 2 * sigma_n;
-else
-	out = 0;
-end

### 2014-04-30 code/covar_hyperparams/exponential_cov_grad.m

@@ -1,0 +1,88 @@
+function grad = exponential_cov_grad(X, y, theta, K_inv)
+
+alpha = K_inv * y;
+term1 = (alpha * alpha' - K_inv);
+
+grad = zeros(numel(theta), 1);
+[d,N] = size(X);
+
+for h = 1:numel(theta)
+	dK = zeros(N,N);
+	for i = 1:N
+		for j = i:N
+			if h <= d
+                if (i == j)
+                    % So, add a tiny bit of noise to X_i.
+                    % Not sure that this is legit, really.
+                    x_temp = X(:,j)+1e-6.*randn(d,1);
+                    dK(j,i) = dk_func_ll(X(:,i), x_temp, theta, h);
+                else
+                    dK(j,i) = dk_func_ll(X(:,i), X(:,j), theta, h);
+                end
+			elseif h == d + 1
+				dK(j,i) = dk_func_sigma_f(X(:,i), X(:,j), theta);
+			elseif h == d + 2
+				dK(j,i) = dk_func_sigma_n(theta, (i==j) );
+			end
+			dK(i, j) = dK(j, i);
+		end
+	end
+	grad(h) = 0.5 * trace(term1*dK);  % Derivative of marginal likelihood 
+end
+
+
+
+
+function out = dk_func_ll(x1, x2, theta, j)
+% Derivative of Matern covariance function with respect to l[j] hyperparameter
+%    x1 - first coordinate
+%    x2 - second coordinate
+%    theta - the array of coordinates that can be changed.
+%       elements 1:d :  l, a vector of scaling params
+%       element d+1: var_f (= sigma_f^2)
+%       element d+2: var_n (= sigma_n^2)
+%    j - hyperparameter
+
+D = numel(x1);
+
+d = x1 - x2;
+u = sqrt(d' * diag(1 ./ (theta(1:D) .^2)) * d);
+
+% d/dl sigma_f * exp(-u) = sigma_f * exp(u) * (-1/(2 * sqrt(sum(d^2/ll^2)))) * (-2 d_j^2 / l_j^3)
+
+out = theta(D+1) * (exp(-u) / u) * (d(j)^2 / theta(j)^3);
+
+
+
+
+function out = dk_func_f(x1, x2, theta)
+% Derivative of Matern covariance function with respect to sigma_f hyperparameter
+%    x1 - first coordinate
+%    x2 - second coordinate
+%    theta - the array of coordinates that can be changed.
[…]

### 2014-04-30 code/covar_hyperparams/exponential_cov_obj.m

@@ -1,0 +1,59 @@
+function obj = exponential_cov_obj( X, y, theta)
+% Negative Marginal likelihood of exponential covariance (objective function to be MINimized)
+
+K = K_mat(X, theta);
+Kinv = pinv(K);
+
+[d, N] = size(X);
+obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
+
+%grad = exponential_cov_grad(X, y, theta, Kinv);
+
+% Use negative of obj and grad, since we have to minimize
+obj = - obj;
+%grad = -1 .* grad;
+
+
+
+
+function out = K_mat(X, theta)
+% Matern covariance function matrix (objective function)
+%    X - data matrix
+%    theta - the struct of coordinates that can be messed with.
+%       Contains ll, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
+
+[d, N] = size(X);
+out = zeros(N);
+for i = 1:N
+	out(i,i) =  k_func(X(:,i), X(:,i), theta, true);
+	for j = i+1:N
+		out(i,j) = k_func(X(:,i), X(:,j), theta, false);
+		out(j,i) = out(i,j);
+	end
+end
+
+
+
+
+function k = k_func(x1, x2, theta, on_diagonal)
+% Matern covariance function
+%    x1 - first coordinate
+%    x2 - second coordinate
+%    theta - the array of coordinates that can be changed.
+%       elements 1:d :  l, a vector of scaling params
+%       element d+1: var_f (= sigma_f^2)
+%       element d+2: var_n (= sigma_n^2)
+%    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
+
+% For now, assume nu = 1/2
+
+D = numel(x1);  % dimensions
+
+% k = theta.var_f * exp(-norm(x1 - x2)/ theta.l) + theta.var_n * [1 if on diagonal; 0 otherwise]
+if (on_diagonal == false)
+	d = x1 - x2;
+	k = theta(D+1) * exp( - sqrt( d' * diag(1 ./ (theta(1:D) .^2)) * d ));
+else
+	% exp(0) = 1, so
+	k = theta(D+1) + theta(D+2);
+end

### 2014-04-30 code/covar_hyperparams/k_func.m

@@ -1,22 +1,0 @@
-function k = k_func(x1, x2, theta, on_diagonal)
-% Matern covariance function
-%    x1 - first coordinate
-%    x2 - second coordinate
-%    theta - the array of coordinates that can be changed.
-%       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
-%    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
-
-% For now, assume nu = 1/2
-
-D = numel(x1);  % dimensions
-
-% k = theta.var_f * exp(-norm(x1 - x2)/ theta.l) + theta.var_n * [1 if on diagonal; 0 otherwise]
-if (on_diagonal == false)
-	d = x1 - x2;
-	k = theta(D+1) * exp( - sqrt( d' * diag(1 ./ (theta(1:D) .^2)) * d ));
-else
-	% exp(0) = 1, so
-	k = theta(D+1) + theta(D+2);
-end

### 2014-04-30 code/covar_hyperparams/k_grad.m

@@ -1,31 +1,0 @@
-function grad = k_grad(X, y, theta, K_inv)
-
-alpha = K_inv * y;
-term1 = (alpha * alpha' - K_inv);
-
-grad = zeros(numel(theta), 1);
-[d,N] = size(X);
-
-for h = 1:numel(theta)
-	dK = zeros(N,N);
-	for i = 1:N
-		for j = i:N
-			if h <= d
-                if (i == j)
-                    % So, add a tiny bit of noise to X_i.
-                    % Not sure that this is legit, really.
-                    x_temp = X(:,j)+1e-6.*randn(d,1);
-                    dK(j,i) = dk_func_ll(X(:,i), x_temp, theta, h);
-                else
-                    dK(j,i) = dk_func_ll(X(:,i), X(:,j), theta, h);
-                end
-			elseif h == d + 1
-				dK(j,i) = dk_func_sigma_f(X(:,i), X(:,j), theta);
-			elseif h == d + 2
-				dK(j,i) = dk_func_sigma_n(theta, (i==j) );
-			end
-			dK(i, j) = dK(j, i);
-		end
-	end
-	grad(h) = 0.5 * trace(term1*dK);  % Derivative of marginal likelihood 
-end

### 2014-04-30 code/covar_hyperparams/k_obj.m

@@ -1,14 +1,0 @@
-function obj = k_obj( X, y, theta)
-% Negative Marginal likelihood (objective function to be MINimized)
-
-K = K_mat(X, theta);
-Kinv = pinv(K);
-
-[d, N] = size(X);
-obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
-
-%grad = k_grad(X, y, theta, Kinv);
-
-% Use negative of obj and grad, since we have to minimize
-obj = - obj;
-%grad = -1 .* grad;

### 2014-04-30 code/optimize_covar.m

@@ -1,22 +1,0 @@
-function [theta, fval] = optimize_covar(X,y)
-
-addpath covar_hyperparams/.
-
-options = optimoptions(@fmincon,...
-'Display','iter', ...
-'Algorithm','interior-point');  % or trust-region-reflective
-
-n_theta = 29;
-theta_0 = ones(n_theta, 1);
-
-Xt = X'; % Each column of X is a 27-param table entries
-%y is a column vector
-
-objective = @(theta) k_obj(Xt,y,theta);
-
-% Lower bound - bound all to be >= 0
-lb = zeros(n_theta,1);
-
-[theta, fval] = fmincon(objective,theta_0,...
-[],[],[],[],lb,[],[],options);
-

### 2014-04-30 code/optimize_hyperparams.m

@@ -1,0 +1,31 @@
+function [theta, fval] = optimize_hyperparams(X,y,cov_type)
+
+addpath covar_hyperparams/.
+
+if cov_type == 2  % Matern (1/2) [exponential covariance]
+
+	options = optimoptions(@fmincon,...
+	'Display','iter', ...
+	'Algorithm','interior-point');  % or trust-region-reflective
+
+	n_theta = 29;
+	theta_0 = ones(n_theta, 1);
+
+	Xt = X'; % Each column of X is a 27-param table entries
+	%y is a column vector
+
+	objective = @(theta) exponential_cov_obj(Xt,y,theta);
+
+	% Lower bound - bound all to be >= 0
+	lb = zeros(n_theta,1);
+
+	[theta, fval] = fmincon(objective,theta_0,...
+	[],[],[],[],lb,[],[],options);
+
+	% We made the function negative to minimize.  Switch sign back.
+	fval = -fval;
+else
+	error('Wrong covariance type.  This only supports Matern (1/2) right now');
+	theta = ones(27);
+	fval = 0;
+end

### 2014-05-02 code/covar_hyperparams/matern_1p5_cov_grad.m
Add hyperparameter optimizations for other covariance functions
We're going to support 4 covariance models:
* Exponential ( Matern with nu = 1/2 )
* Matern with nu = 3/2
* Matern with nu = 5/2
* Squared exponential

@@ -1,4 +1,5 @@
-function grad = exponential_cov_grad(X, y, theta, K_inv)
+function grad = matern_1p5_cov_grad(X, y, theta, K_inv)
+% Gradient of Matern covariance (nu = 3/2)
 
 alpha = K_inv * y;
 term1 = (alpha * alpha' - K_inv);
@@ -11,14 +12,7 @@ for h = 1:numel(theta)
 	for i = 1:N
 		for j = i:N
 			if h <= d
-				if (i == j)
-					% So, add a tiny bit of noise to X_i.
-					% Not sure that this is legit, really.
-					x_temp = X(:,j)+1e-6.*randn(d,1);
-					dK(j,i) = dk_func_ll(X(:,i), x_temp, theta, h);
-				else
-					dK(j,i) = dk_func_ll(X(:,i), X(:,j), theta, h);
-				end
+				dK(j,i) = dk_func_ll(X(:,i), X(:,j), theta, h);
 			elseif h == d + 1
 				dK(j,i) = dk_func_sigma_f(X(:,i), X(:,j), theta);
 			elseif h == d + 2
@@ -34,7 +28,7 @@ end
 
 
 function out = dk_func_ll(x1, x2, theta, j)
-% Derivative of Matern covariance function with respect to l[j] hyperparameter
+% Derivative of Matern covariance (nu = 3/2) function with respect to l[j] hyperparameter
 %    x1 - first coordinate
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
@@ -46,17 +40,17 @@ function out = dk_func_ll(x1, x2, theta, j)
 D = numel(x1);
 
 d = x1 - x2;
-u = sqrt(d' * diag(1 ./ (theta(1:D) .^2)) * d);
+u = sqrt(3 * d' * diag(1 ./ theta(1:D) .^2) * d);
 
-% d/dl sigma_f * exp(-u) = sigma_f * exp(-u) * -(1/(2 * sqrt(sum(d^2/ll^2)))) * (-2 d_j^2 / l_j^3)
-
-out = theta(D+1) * (exp(-u) / u) * (d(j)^2 / theta(j)^3);
+% d/dl sigma_f * (1 +u) * exp(-u) = sigma_f * -u * exp(-u) * -(sqrt(3)/(2 * sqrt(sum(d^2/ll^2)))) * (-2 d_j^2 / l_j^3)
[…]

### 2014-05-02 code/covar_hyperparams/matern_1p5_cov_obj.m

@@ -1,5 +1,5 @@
-function obj = exponential_cov_obj( X, y, theta)
-% Negative Marginal likelihood of exponential covariance (objective function to be MINimized)
+function [obj, grad] = matern_1p5_cov_obj( X, y, theta)
+% Negative Marginal likelihood of Matern covariance (nu = 3/2) (objective function to be MINimized)
 
 K = K_mat(X, theta);
 try
@@ -11,17 +11,17 @@ end
 [d, N] = size(X);
 obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
 
-%grad = exponential_cov_grad(X, y, theta, Kinv);
+grad = matern_1p5_cov_grad(X, y, theta, Kinv);
 
 % Use negative of obj and grad, since we have to minimize
 obj = - obj;
-%grad = -1 .* grad;
+grad = -1 .* grad;
 
 
 
 
 function out = K_mat(X, theta)
-% Matern covariance function matrix (objective function)
+% Matern covariance function matrix (nu = 3/2) (objective function)
 %    X - data matrix
 %    theta - the struct of coordinates that can be messed with.
 %       Contains ll, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
@@ -40,7 +40,7 @@ end
 
 
 function k = k_func(x1, x2, theta, on_diagonal)
-% Matern covariance function
+% Matern covariance function (nu = 3/2)
 %    x1 - first coordinate
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
@@ -49,15 +49,15 @@ function k = k_func(x1, x2, theta, on_diagonal)
 %       element d+2: var_n (= sigma_n^2)
 %    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
 
-% For now, assume nu = 1/2
-
 D = numel(x1);  % dimensions
 
-% k = theta.var_f * exp(-norm(x1 - x2)/ theta.l) + theta.var_n * [1 if on diagonal; 0 otherwise]
+% u = 1 + sqrt(3)*(x1-x2)/l
+% k = theta.var_f * (1 + u) * exp(-u) + theta.var_n * [1 if on diagonal; 0 otherwise]
 if (on_diagonal == false)
 	d = x1 - x2;
[…]

### 2014-05-02 code/covar_hyperparams/matern_2p5_cov_grad.m

@@ -1,4 +1,5 @@
-function grad = exponential_cov_grad(X, y, theta, K_inv)
+function grad = matern_2p5_cov_grad(X, y, theta, K_inv)
+% Gradient of Matern covariance (nu = 5/2)
 
 alpha = K_inv * y;
 term1 = (alpha * alpha' - K_inv);
@@ -11,14 +12,7 @@ for h = 1:numel(theta)
 	for i = 1:N
 		for j = i:N
 			if h <= d
-				if (i == j)
-					% So, add a tiny bit of noise to X_i.
-					% Not sure that this is legit, really.
-					x_temp = X(:,j)+1e-6.*randn(d,1);
-					dK(j,i) = dk_func_ll(X(:,i), x_temp, theta, h);
-				else
-					dK(j,i) = dk_func_ll(X(:,i), X(:,j), theta, h);
-				end
+				dK(j,i) = dk_func_ll(X(:,i), X(:,j), theta, h);
 			elseif h == d + 1
 				dK(j,i) = dk_func_sigma_f(X(:,i), X(:,j), theta);
 			elseif h == d + 2
@@ -34,7 +28,7 @@ end
 
 
 function out = dk_func_ll(x1, x2, theta, j)
-% Derivative of Matern covariance function with respect to l[j] hyperparameter
+% Derivative of Matern covariance (nu = 5/2) function with respect to l[j] hyperparameter
 %    x1 - first coordinate
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
@@ -46,17 +40,16 @@ function out = dk_func_ll(x1, x2, theta, j)
 D = numel(x1);
 
 d = x1 - x2;
-u = sqrt(d' * diag(1 ./ (theta(1:D) .^2)) * d);
+u = sqrt(5 * d' * diag(1 ./ theta(1:D) .^2) * d);
 
-% d/dl sigma_f * exp(-u) = sigma_f * exp(-u) * -(1/(2 * sqrt(sum(d^2/ll^2)))) * (-2 d_j^2 / l_j^3)
-
-out = theta(D+1) * (exp(-u) / u) * (d(j)^2 / theta(j)^3);
+% d/dl sigma_f * (1 +u + u^2) * exp(-u)
+out = 5 * theta(D+1) * (1 - u) * exp(-u) * (d(j)^2 / theta(j)^3);
 
 
 
 
 function out = dk_func_sigma_f(x1, x2, theta)
-% Derivative of Matern covariance function with respect to sigma_f hyperparameter
[…]

### 2014-05-02 code/covar_hyperparams/matern_2p5_cov_obj.m

@@ -1,5 +1,5 @@
-function obj = exponential_cov_obj( X, y, theta)
-% Negative Marginal likelihood of exponential covariance (objective function to be MINimized)
+function [obj, grad] = matern_2p5_cov_obj( X, y, theta)
+% Negative Marginal likelihood of Matern covariance (nu = 5/2) (objective function to be MINimized)
 
 K = K_mat(X, theta);
 try
@@ -11,17 +11,17 @@ end
 [d, N] = size(X);
 obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
 
-%grad = exponential_cov_grad(X, y, theta, Kinv);
+grad = matern_2p5_cov_grad(X, y, theta, Kinv);
 
 % Use negative of obj and grad, since we have to minimize
 obj = - obj;
-%grad = -1 .* grad;
+grad = -1 .* grad;
 
 
 
 
 function out = K_mat(X, theta)
-% Matern covariance function matrix (objective function)
+% Matern covariance function matrix (nu = 5/2) (objective function)
 %    X - data matrix
 %    theta - the struct of coordinates that can be messed with.
 %       Contains ll, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
@@ -40,7 +40,7 @@ end
 
 
 function k = k_func(x1, x2, theta, on_diagonal)
-% Matern covariance function
+% Matern covariance function (nu = 5/2)
 %    x1 - first coordinate
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
@@ -49,15 +49,15 @@ function k = k_func(x1, x2, theta, on_diagonal)
 %       element d+2: var_n (= sigma_n^2)
 %    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
 
-% For now, assume nu = 1/2
-
 D = numel(x1);  % dimensions
 
-% k = theta.var_f * exp(-norm(x1 - x2)/ theta.l) + theta.var_n * [1 if on diagonal; 0 otherwise]
+% u = 1 + sqrt(3)*(x1-x2)/l
+% k = theta.var_f * (1 + u) * exp(-u) + theta.var_n * [1 if on diagonal; 0 otherwise]
 if (on_diagonal == false)
 	d = x1 - x2;
[…]

### 2014-05-02 code/covar_hyperparams/se_cov_grad.m

@@ -1,4 +1,4 @@
-function grad = exponential_cov_grad(X, y, theta, K_inv)
+function grad = se_cov_grad(X, y, theta, K_inv)
 
 alpha = K_inv * y;
 term1 = (alpha * alpha' - K_inv);
@@ -11,14 +11,7 @@ for h = 1:numel(theta)
 	for i = 1:N
 		for j = i:N
 			if h <= d
-				if (i == j)
-					% So, add a tiny bit of noise to X_i.
-					% Not sure that this is legit, really.
-					x_temp = X(:,j)+1e-6.*randn(d,1);
-					dK(j,i) = dk_func_ll(X(:,i), x_temp, theta, h);
-				else
-					dK(j,i) = dk_func_ll(X(:,i), X(:,j), theta, h);
-				end
+				dK(j,i) = dk_func_ll(X(:,i), X(:,j), theta, h);
 			elseif h == d + 1
 				dK(j,i) = dk_func_sigma_f(X(:,i), X(:,j), theta);
 			elseif h == d + 2
@@ -34,7 +27,7 @@ end
 
 
 function out = dk_func_ll(x1, x2, theta, j)
-% Derivative of Matern covariance function with respect to l[j] hyperparameter
+% Derivative of SE covariance function with respect to l[j] hyperparameter
 %    x1 - first coordinate
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
@@ -46,17 +39,17 @@ function out = dk_func_ll(x1, x2, theta, j)
 D = numel(x1);
 
 d = x1 - x2;
-u = sqrt(d' * diag(1 ./ (theta(1:D) .^2)) * d);
+u = d' * diag(1 ./ (theta(1:D) .^2)) * d;
 
-% d/dl sigma_f * exp(-u) = sigma_f * exp(-u) * -(1/(2 * sqrt(sum(d^2/ll^2)))) * (-2 d_j^2 / l_j^3)
+% d/dl sigma_f * exp(-u) = sigma_f * exp(-u) * -(-2 d_j^2 / l_j^3)
 
-out = theta(D+1) * (exp(-u) / u) * (d(j)^2 / theta(j)^3);
+out = 2 * theta(D+1) * exp(-u) * (d(j)^2 / theta(j)^3);
 
 
 
 
 function out = dk_func_sigma_f(x1, x2, theta)
-% Derivative of Matern covariance function with respect to sigma_f hyperparameter
[…]

### 2014-05-02 code/covar_hyperparams/se_cov_obj.m

@@ -1,5 +1,5 @@
-function obj = exponential_cov_obj( X, y, theta)
-% Negative Marginal likelihood of exponential covariance (objective function to be MINimized)
+function [obj, grad] = se_cov_obj( X, y, theta)
+% Negative Marginal likelihood of squared exponential covariance (SE) (objective function to be MINimized)
 
 K = K_mat(X, theta);
 try
@@ -11,17 +11,17 @@ end
 [d, N] = size(X);
 obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
 
-%grad = exponential_cov_grad(X, y, theta, Kinv);
+grad = se_cov_grad(X, y, theta, Kinv);
 
 % Use negative of obj and grad, since we have to minimize
 obj = - obj;
-%grad = -1 .* grad;
+grad = -1 .* grad;
 
 
 
 
 function out = K_mat(X, theta)
-% Matern covariance function matrix (objective function)
+% SE covariance function matrix (objective function)
 %    X - data matrix
 %    theta - the struct of coordinates that can be messed with.
 %       Contains ll, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
@@ -40,7 +40,7 @@ end
 
 
 function k = k_func(x1, x2, theta, on_diagonal)
-% Matern covariance function
+% SE covariance function
 %    x1 - first coordinate
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
@@ -56,7 +56,7 @@ D = numel(x1);  % dimensions
 % k = theta.var_f * exp(-norm(x1 - x2)/ theta.l) + theta.var_n * [1 if on diagonal; 0 otherwise]
 if (on_diagonal == false)
 	d = x1 - x2;
-	k = theta(D+1) * exp( - sqrt( d' * diag(1 ./ (theta(1:D) .^2)) * d ));
+	k = theta(D+1) * exp( - d' * diag(1 ./ (theta(1:D) .^2)) * d );
 else
 	% exp(0) = 1, so
 	k = theta(D+1) + theta(D+2);

### 2014-05-02 code/optimize_hyperparams.m

@@ -2,30 +2,47 @@ function [theta, fval] = optimize_hyperparams(X,y,cov_type)
 
 addpath covar_hyperparams/.
 
-if cov_type == 2  % Matern (1/2) [exponential covariance]
+n_theta = 29;
+theta_0 = ones(n_theta, 1);
 
-	options = optimoptions(@fmincon,...
-	'Display','iter', ...
-	'Algorithm','interior-point');  % or trust-region-reflective
+Xt = X'; % Each column of X is a 27-param table entries
+%y is a column vector
 
-	n_theta = 29;
-	theta_0 = ones(n_theta, 1);
+% Lower bound - bound all to be >= 0
+lb = zeros(n_theta,1);
 
-	Xt = X'; % Each column of X is a 27-param table entries
-	%y is a column vector
-
-	objective = @(theta) exponential_cov_obj(Xt,y,theta);
-
-	% Lower bound - bound all to be >= 0
-	lb = zeros(n_theta,1);
-
-	[theta, fval] = fmincon(objective,theta_0,...
-	[],[],[],[],lb,[],[],options);
-
-	% We made the function negative to minimize.  Switch sign back.
-	fval = -fval;
-else
-	error('Wrong covariance type.  This only supports Matern (1/2) right now');
-	theta = ones(27);
-	fval = 0;
+switch cov_type
+	case 1  % Squared exponential
+		options = optimoptions(@fmincon,...
+			'Display','iter', ...
+			'GradObj','on', ...
+			'Algorithm','interior-point'); %trust-region-reflective');
+		objective = @(theta) se_cov_obj(Xt,y,theta);
+	case 2  % Matern (1/2) [exponential covariance]
+		options = optimoptions(@fmincon,...
+			'Display','iter', ...
+			'Algorithm','interior-point');  % or trust-region-reflective
+		objective = @(theta) exponential_cov_obj(Xt,y,theta);
+	case 3  % Matern (3/2)
+		options = optimoptions(@fmincon,...
+			'Display','iter', ...
+			'GradObj','on', ...
+			'Algorithm', 'interior-point'); %'trust-region-reflective');
+		objective = @(theta) matern_1p5_cov_obj(Xt,y,theta);
+	case 4  % Matern (5/2)
+		options = optimoptions(@fmincon,...
+			'Display','iter', ...
+			'GradObj','on', ...
+			'Algorithm', 'interior-point'); %'trust-region-reflective');
[…]

### 2014-05-02 code/run_optimization.m

@@ -1,0 +1,46 @@
+%function [theta, y_val] = run_optimization()
+
+%Call readProcessedData to get our data from the file into y and X
+[y, X] = readProcessedData();
+
+% Scale data in the same way as GPR code
+mean_y = mean(y);
+stdev_y = std(y);
+mean_X = mean(X,1);
+stdev_X = std(X,0,1);
+Scaled_X = zeros(size(X));
+Scaled_y = y - mean_y;
+N = size(X, 1);
+for i = 1:N
+	Scaled_X(i,:) = (X(i,:) - mean_X) ./ stdev_X;
+end
+
+% Run the optimizations several times
+D = size(X, 2);
+%theta = zeros(D+2,4,4);
+%fval = zeros(4,4);
+
+for cov_type = 2:4
+ 	[theta(:, cov_type, 1), fval(cov_type, 1)] = optimize_hyperparams( Scaled_X(1:100,:), Scaled_y(1:100), cov_type);
+ 	[theta(:, cov_type, 2), fval(cov_type, 2)] = optimize_hyperparams( Scaled_X(101:200,:), Scaled_y(101:200), cov_type);
+ 	randSet = randperm(nData, 100);
+ 	[theta(:, cov_type, 3), fval(cov_type, 3)] = optimize_hyperparams( Scaled_X(randSet,:), Scaled_y(randSet), cov_type);
+
+	% Also - check to see if this makes things better.  (Scale the sample separately.)
+	% This would mean that we would scale 1-10000, then rescale at each time step.
+	SampleSize = 100;
+	Scaled_X = zeros(SampleSize, D);
+	mean_y = mean(y(1:SampleSize));
+	stdev_y = std(y(1:SampleSize));
+	mean_X = mean(X(1:SampleSize, :),1);
+	stdev_X = std(X(1:SampleSize, :),0,1);
+	Scaled_y = y - mean_y;
+	for i = 1:N
+		Scaled_X(i,:) = (X(i,:) - mean_X) ./ stdev_X;
+	end
+	[theta(:, cov_type, 4), fval(cov_type, 4)] = optimize_hyperparams( Scaled_X(1:100,:), Scaled_y(1:100), cov_type);
+
+end
+
+save('optimized_theta', 'theta')
+save('optimized_fval', 'fval')

### 2014-05-03 code/Gaussian_Processes_Regression.m
Modified Code

@@ -42,16 +42,21 @@ Design_X = Scaled_X'; Design_y = Scaled_y';
 X_Window = Design_X(:,1:windowSize); y_Window = Design_y(1:windowSize);
 ctr = windowSize;
 
-Predicted_Data = [];
-while (ctr < size(Design_X,2))
-    
-    %Select the hyper-parameters
-    sigma_f = 1.0; sigma_n = 0.1;
-    l = ones(size(X_Window,1),1)';   %All characteristic length scales are initialized to 1
+%Select the hyper-parameters
+sigma_f = 1.0; sigma_n = 0.1;
+l = ones(size(X_Window,1),1)';   %All characteristic length scales are initialized to 1
 
-    %Build the Gram matrix, K
-    num_examples = size(X_Window,2);
-    K = ones(num_examples);
+%Check to see whether K is stored. If not, build the matrix from scratch
+num_examples = size(X_Window,2);
+K = zeros(num_examples);
+fileName = 'K_Matrix';
+fileName = strcat(fileName, '_', num2str(windowSize),'_', num2str(cov_type), '.txt');
+path = '../data/';
+if exist([path fileName], 'file')
+   K = dlmread([path fileName],'\t');
+else
+    %Build the Gram matrix - K, using all of the data in X_Window
+    fprintf('Building the initial Gram matrix - K using %d examples \n', windowSize);        
     for i=1:num_examples
         for j=i:num_examples  
             switch cov_type
@@ -72,9 +77,47 @@ while (ctr < size(Design_X,2))
                 K(j,i) = K(i,j);  %K should be symmetric
             end
         end
+        if (mod(i,100) == 0)
+            fprintf('Finished building row no. %d\n', i);
+        end
     end
+    writeFile(K, fileName);
+end
 
-    %Train the hyperparameters
+
+%Train the hyperparameters of the initial K matrix
+
+
+Predicted_yValues = [];
+while (ctr < size(Design_X,2))        
+    
+    %Build the delta Gram matrix
+    if (ctr ~= windowSize)
+        K(1,:) = []; K(:,1) = [];
+        new_X = X_Window(:,end);
+        k_delta = ones(num_examples,1);
+        for i=1:num_examples  
+            switch cov_type
+                case 1
+                    k_delta(i) = Squared_Exponential(X_Window(:,i), new_X, sigma_f, l);
+                case 2
+                    k_delta(i) = Ornstein_Uhlenbeck(X_Window(:,i), new_X, sigma_f, l);
[…]

### 2014-05-03 code/convert_to_regdata_GPR.py

@@ -1,0 +1,79 @@
+#!/usr/bin/python
+
+# This file assumes that timestamps are always increasing
+import os
+import sys
+
+def __main__():
+
+    if len ( sys.argv ) < 3 :
+        print ("Usage: yhat-file msecs_diff [trainingSampleSize] [testSampleSize]")
+        sys.exit()
+
+    # expect format of yhat-file: 
+    # MFM-disc TARGETPRICE BIDPX BIDSZ ASKPX ASKSZ VAR1 ...
+    # 39706307 4779.455624 4779. 50.00 4780. 40.00 ...
+
+    msecs_diff = int(sys.argv[2])
+    if len ( sys.argv ) == 5 :
+    	train_sample_size = int(sys.argv[3])
+    	test_sample_size = int(sys.argv[4])
+    else :
+	train_sample_size = -100
+    	test_sample_size = -100
+    line_vec_ = []
+
+    # Initialize the file handles for the test file and targetprice file
+    path = "../data"    
+    outputfilename_ = "prod_data_origprice.txt"
+    pathname_outputfile = os.path.join(path, outputfilename_)
+    output_target_file_handle_ = open ( pathname_outputfile, 'w+') # open file for writing the targetprice file
+    testfilename_ = "prod_test_data.txt"	# test file for comparing GPReg with linear regression / pca regression
+    pathname_testfile = os.path.join(path, testfilename_)
+    test_file_handle_ = open ( pathname_testfile, 'w+') # open file for writing
+
+    num_lines = 0
+    with open(sys.argv[1]) as input_yhat_file_handle_ :		#argv[1] is prod_data_v.txt
+        for input_yhat_line_ in input_yhat_file_handle_ :
+            input_yhat_line_words_ = input_yhat_line_.strip().split()
+            if len ( input_yhat_line_words_ ) >= 6 :
+                timeval_ = (int)(input_yhat_line_words_[0])
+                tgtpx_ = (float)(input_yhat_line_words_[1])
+                #                bidpx_ = (float)(input_yhat_line_words_[2])
+                #                bidsz_ = (float)(input_yhat_line_words_[3])
+                #                askpx_ = (float)(input_yhat_line_words_[4])
+                #                asksz_ = (float)(input_yhat_line_words_[5])
+                
+                line_vec_.append ( input_yhat_line_ )
+                oldest_yhat_line_ = line_vec_[0]
+                oldest_yhat_line_words_ = oldest_yhat_line_.strip().split()
+
[…]

### 2014-05-03 code/covar_hyperparams/exponential_cov_grad.m
Fixes to covariance functions

@@ -39,8 +39,8 @@ function out = dk_func_ll(x1, x2, theta, j)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 %    j - hyperparameter
 
 D = numel(x1);
@@ -50,7 +50,7 @@ u = sqrt(d' * diag(1 ./ (theta(1:D) .^2)) * d);
 
 % d/dl sigma_f * exp(-u) = sigma_f * exp(-u) * -(1/(2 * sqrt(sum(d^2/ll^2)))) * (-2 d_j^2 / l_j^3)
 
-out = theta(D+1) * (exp(-u) / u) * (d(j)^2 / theta(j)^3);
+out = theta(D+1)^2 * (exp(-u) / u) * (d(j)^2 / theta(j)^3);
 
 
 
@@ -61,12 +61,12 @@ function out = dk_func_sigma_f(x1, x2, theta)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 
 D = numel(x1);
 d = x1 - x2;
-sigma_f = sqrt(theta(D+1));
+sigma_f = theta(D+1);
 out = 2 * sigma_f * exp( - sqrt( d' * diag(1 ./ (theta(1:D) .^2)) * d ));
 
 
@@ -76,12 +76,12 @@ function out = dk_func_sigma_n(theta, on_diagonal)
 % Derivative of Matern covariance function with respect to sigma_f hyperparameter
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 %    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
 
 if (on_diagonal == true)
-	sigma_n = sqrt(theta(end));
+	sigma_n = theta(end);
 	out = 2 * sigma_n;
 else
 	out = 0;

### 2014-05-03 code/covar_hyperparams/exponential_cov_obj.m

@@ -9,7 +9,8 @@ catch exception
 end
 
 [d, N] = size(X);
-obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
+% sum(log(svd(K))) is equivalent to log(det(K)), but less likely to blow up
+obj = - 0.5 * y' * Kinv * y - 0.5 * sum(log(svd(K))) - N/2 * log(2 * pi);
 
 %grad = exponential_cov_grad(X, y, theta, Kinv);
 
@@ -24,7 +25,7 @@ function out = K_mat(X, theta)
 % Matern covariance function matrix (objective function)
 %    X - data matrix
 %    theta - the struct of coordinates that can be messed with.
-%       Contains ll, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
+%       Contains ll, a vector of scaling params, and sigma_f and sigma_n
 
 [d, N] = size(X);
 out = zeros(N);
@@ -45,8 +46,8 @@ function k = k_func(x1, x2, theta, on_diagonal)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 %    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
 
 % For now, assume nu = 1/2
@@ -56,8 +57,8 @@ D = numel(x1);  % dimensions
 % k = theta.var_f * exp(-norm(x1 - x2)/ theta.l) + theta.var_n * [1 if on diagonal; 0 otherwise]
 if (on_diagonal == false)
 	d = x1 - x2;
-	k = theta(D+1) * exp( - sqrt( d' * diag(1 ./ (theta(1:D) .^2)) * d ));
+	k = theta(D+1)^2 * exp( - sqrt( d' * diag(1 ./ (theta(1:D) .^2)) * d ));
 else
 	% exp(0) = 1, so
-	k = theta(D+1) + theta(D+2);
+	k = theta(D+1)^2 + theta(D+2)^2;
 end

### 2014-05-03 code/covar_hyperparams/matern_1p5_cov_grad.m

@@ -33,8 +33,8 @@ function out = dk_func_ll(x1, x2, theta, j)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 %    j - hyperparameter
 
 D = numel(x1);
@@ -44,7 +44,7 @@ u = sqrt(3 * d' * diag(1 ./ theta(1:D) .^2) * d);
 
 % d/dl sigma_f * (1 +u) * exp(-u) = sigma_f * -u * exp(-u) * -(sqrt(3)/(2 * sqrt(sum(d^2/ll^2)))) * (-2 d_j^2 / l_j^3)
 %          = sigma_f * exp(-u) * (3 d_j^2 / l_j^3)
-out = 3 * theta(D+1) * exp(-u) * (d(j)^2 / theta(j)^3);
+out = 3 * theta(D+1)^2 * exp(-u) * (d(j)^2 / theta(j)^3);
 
 
 
@@ -55,11 +55,11 @@ function out = dk_func_sigma_f(x1, x2, theta)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 
 D = numel(x1);
-sigma_f = sqrt(theta(D+1));
+sigma_f = theta(D+1);
 d = x1 - x2;
 u = sqrt(3 * d' * diag(1 ./ theta(1:D) .^2) * d);
 out = 2 * sigma_f * (1 + u) * exp(-u);
@@ -71,12 +71,12 @@ function out = dk_func_sigma_n(theta, on_diagonal)
 % Derivative of Matern covariance function (nu = 3/2) with respect to sigma_f hyperparameter
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
[…]

### 2014-05-03 code/covar_hyperparams/matern_1p5_cov_obj.m

@@ -9,7 +9,8 @@ catch exception
 end
 
 [d, N] = size(X);
-obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
+% sum(log(svd(K))) is equivalent to log(det(K)), but less likely to blow up
+obj = - 0.5 * y' * Kinv * y - 0.5 * sum(log(svd(K))) - N/2 * log(2 * pi);
 
 grad = matern_1p5_cov_grad(X, y, theta, Kinv);
 
@@ -24,7 +25,7 @@ function out = K_mat(X, theta)
 % Matern covariance function matrix (nu = 3/2) (objective function)
 %    X - data matrix
 %    theta - the struct of coordinates that can be messed with.
-%       Contains ll, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
+%       Contains ll, a vector of scaling params, and sigma_f and sigma_n
 
 [d, N] = size(X);
 out = zeros(N);
@@ -45,8 +46,8 @@ function k = k_func(x1, x2, theta, on_diagonal)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 %    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
 
 D = numel(x1);  % dimensions
@@ -56,8 +57,8 @@ D = numel(x1);  % dimensions
 if (on_diagonal == false)
 	d = x1 - x2;
 	u = sqrt(3 * d' * diag(1 ./ theta(1:D) .^2) * d);
-	k = theta(D+1) * (1 + u) * exp(-u);
+	k = theta(D+1)^2 * (1 + u) * exp(-u);
 else
 	% u = 0, and (1 + u) exp(-u) = (1) * exp(0) = 1, so
-	k = theta(D+1) + theta(D+2);
+	k = theta(D+1)^2 + theta(D+2)^2;
 end

### 2014-05-03 code/covar_hyperparams/matern_2p5_cov_grad.m

@@ -33,8 +33,8 @@ function out = dk_func_ll(x1, x2, theta, j)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 %    j - hyperparameter
 
 D = numel(x1);
@@ -43,7 +43,7 @@ d = x1 - x2;
 u = sqrt(5 * d' * diag(1 ./ theta(1:D) .^2) * d);
 
 % d/dl sigma_f * (1 +u + u^2) * exp(-u)
-out = 5 * theta(D+1) * (1 - u) * exp(-u) * (d(j)^2 / theta(j)^3);
+out = 5 * theta(D+1)^2 * (1 - u) * exp(-u) * (d(j)^2 / theta(j)^3);
 
 
 
@@ -54,11 +54,11 @@ function out = dk_func_sigma_f(x1, x2, theta)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 
 D = numel(x1);
-sigma_f = sqrt(theta(D+1));
+sigma_f = theta(D+1);
 d = x1 - x2;
 u = sqrt(5 * d' * diag(1 ./ theta(1:D) .^2) * d);
 out = 2 * sigma_f * (1 + u + u^2) * exp(-u);
@@ -70,12 +70,12 @@ function out = dk_func_sigma_n(theta, on_diagonal)
 % Derivative of Matern covariance function (nu = 5/2) with respect to sigma_f hyperparameter
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 %    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
 
 if (on_diagonal == true)
-	sigma_n = sqrt(theta(end));
+	sigma_n = theta(end);
 	out = 2 * sigma_n;
 else
 	out = 0;

### 2014-05-03 code/covar_hyperparams/matern_2p5_cov_obj.m

@@ -9,7 +9,8 @@ catch exception
 end
 
 [d, N] = size(X);
-obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
+% sum(log(svd(K))) is equivalent to log(det(K)), but less likely to blow up
+obj = - 0.5 * y' * Kinv * y - 0.5 * sum(log(svd(K))) - N/2 * log(2 * pi);
 
 grad = matern_2p5_cov_grad(X, y, theta, Kinv);
 
@@ -24,7 +25,7 @@ function out = K_mat(X, theta)
 % Matern covariance function matrix (nu = 5/2) (objective function)
 %    X - data matrix
 %    theta - the struct of coordinates that can be messed with.
-%       Contains ll, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
+%       Contains ll, a vector of scaling params, and sigma_f and sigma_n
 
 [d, N] = size(X);
 out = zeros(N);
@@ -45,8 +46,8 @@ function k = k_func(x1, x2, theta, on_diagonal)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 %    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
 
 D = numel(x1);  % dimensions
@@ -56,8 +57,8 @@ D = numel(x1);  % dimensions
 if (on_diagonal == false)
 	d = x1 - x2;
 	u = sqrt(5 * d' * diag(1 ./ theta(1:D) .^2) * d);
-	k = theta(D+1) * (1 + u + u^2) * exp(-u);
+	k = theta(D+1)^2 * (1 + u + u^2) * exp(-u);
 else
 	% u = 0, and (1 + u + u^2) exp(-u) = (1) * exp(0) = 1, so
-	k = theta(D+1) + theta(D+2);
+	k = theta(D+1)^2 + theta(D+2)^2;
 end

### 2014-05-03 code/covar_hyperparams/se_cov_grad.m

@@ -32,18 +32,18 @@ function out = dk_func_ll(x1, x2, theta, j)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 %    j - hyperparameter
 
 D = numel(x1);
 
 d = x1 - x2;
-u = d' * diag(1 ./ (theta(1:D) .^2)) * d;
+u = 0.5 * d' * diag(1 ./ (theta(1:D) .^2)) * d;
 
-% d/dl sigma_f * exp(-u) = sigma_f * exp(-u) * -(-2 d_j^2 / l_j^3)
+% d/dl sigma_f^2 * exp(-u) = sigma_f^2 * exp(-u) * -1/2 * (-2 d_j^2 / l_j^3)
 
-out = 2 * theta(D+1) * exp(-u) * (d(j)^2 / theta(j)^3);
+out = theta(D+1)^2 * exp(-u) * (d(j)^2 / theta(j)^3);
 
 
 
@@ -54,13 +54,13 @@ function out = dk_func_sigma_f(x1, x2, theta)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 
 D = numel(x1);
 d = x1 - x2;
-sigma_f = sqrt(theta(D+1));
-out = 2 * sigma_f * exp( - d' * diag(1 ./ (theta(1:D) .^2)) * d );
+sigma_f = theta(D+1);
+out = 2 * sigma_f * exp( - 0.5 * d' * diag(1 ./ (theta(1:D) .^2)) * d );
 
 
 
@@ -69,12 +69,12 @@ function out = dk_func_sigma_n(theta, on_diagonal)
 % Derivative of SE covariance function with respect to sigma_f hyperparameter
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
[…]

### 2014-05-03 code/covar_hyperparams/se_cov_obj.m

@@ -9,8 +9,9 @@ catch exception
 end
 
 [d, N] = size(X);
-obj = - 0.5 * y' * Kinv * y - 0.5 * log(det(K)) - N/2 * log(2 * pi);
-
+% sum(log(svd(K))) is equivalent to log(det(K)), but less likely to blow up
+obj = - 0.5 * y' * Kinv * y - 0.5 * sum(log(svd(K))) - N/2 * log(2 * pi);
+%fprintf('term 1: %f, term 2: %f, term 3: %f\n', -0.5*y'*Kinv*y, -0.5*sum(log(svd(K))), -N/2*log(2*pi));
 grad = se_cov_grad(X, y, theta, Kinv);
 
 % Use negative of obj and grad, since we have to minimize
@@ -24,7 +25,7 @@ function out = K_mat(X, theta)
 % SE covariance function matrix (objective function)
 %    X - data matrix
 %    theta - the struct of coordinates that can be messed with.
-%       Contains ll, a vector of scaling params, and var_f and var_n, the squared sigma's for those values
+%       Contains ll, a vector of scaling params, and sigma_f and sigma_n
 
 [d, N] = size(X);
 out = zeros(N);
@@ -45,19 +46,15 @@ function k = k_func(x1, x2, theta, on_diagonal)
 %    x2 - second coordinate
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
-%       element d+1: var_f (= sigma_f^2)
-%       element d+2: var_n (= sigma_n^2)
+%       element d+1: sigma_f
+%       element d+2: sigma_n
 %    on_diagonal - true if x1 = x2 and this element is on the diagonal of the K matrix.
-
-% For now, assume nu = 1/2
 
 D = numel(x1);  % dimensions
 
-% k = theta.var_f * exp(-norm(x1 - x2)/ theta.l) + theta.var_n * [1 if on diagonal; 0 otherwise]
-if (on_diagonal == false)
-	d = x1 - x2;
-	k = theta(D+1) * exp( - d' * diag(1 ./ (theta(1:D) .^2)) * d );
-else
-	% exp(0) = 1, so
-	k = theta(D+1) + theta(D+2);
+% k = theta.var_f * exp(- 1/2 * (x1 - x2)^2/ theta.l^2) + theta.var_n * [1 if on diagonal; 0 otherwise]
[…]

### 2014-05-03 paper/paper.tex
Add to paper description of hyperparameter optimization

@@
 \usepackage{url}
+\usepackage{amsmath,amssymb}
 %\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
@@
 Madison, WI 53705 \\
-\texttt{@cs.wisc.edu} \\
+\texttt{sde@cs.wisc.edu} \\
+}
 
@@
 %\nipsfinalcopy % Uncomment for camera-ready version
+
+\def\bfx{\mathbf x}
+\def\bfy{\mathbf y}
+\def\bftheta{\mathbf \theta}
+\def\bfl{\mathbf \mathcal l}
+\def\l{\mathcal l}
+\def\R{\mathbb R}
+\def\diag{\mbox{ diag }}
 
@@
 
+% TODO:  Insert a paragraph about normalization of data.
+
+\section{Covariance Functions}
+We consider 4 different covariance functions, all of which can be parameterized in terms of the hyperparameters $\bftheta = \{\bfl, \sigma_n^2 \sigma_f^2\}$:
+Let
+\begin{equation}
+	u = \left( \left( \bfx_p - \bfx_q \right)^\top \diag(\bfl)^{-2} \left( \bfx_p - \bfx_q \right) \right)^{1/2}
+\end{equation}
+
+Then, the covariance functions are:
+\begin{align}
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- \frac{1}{2} u^2} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Squared Exponential}   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Exponential}   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{3} u \right) \exp{- sqrt{3}  u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{3}{2}$) }   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{5} u \right) + 5 u^2) \exp{- \sqrt{5} u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{5}{2}$) }   \\
+\end{align}
+
+$\sigma_n^2$ and $\sigma_f^2$ are the noise and signal level parameters, respectively, and $\l_1, \dots, \l_D$ are characteristic length-scales.
+The covariance matrix $K_y$ is given by $[K_y]_{pq} = k(\bfx_p, \bfx_q)$.
+We can think of these length-scales as a measure of the irrelevance of the associated feature; a large value of $\l_i$ means that the $i$-th parameter of the feature vector has very little effect on the resultant covariance matrix.
+This is a type of automatic relevance determination (ARD).
+
+\subsection{Tuning the hyperparameters}
+We tuned the hyperparameters by maximizing the log marginal likelihood, 
+\begin{equation}
+	\log p(\bfy | X, \bftheta) = - \frac{1}{2} \bfy^\top K_y^{-1} \bfy - \frac{1}{2} \log \| K_y \| - \frac{n}{2} \log (2 \pi)
+\end{equation}
+
+Russell and Williams state that the maximum of this quantity is generally more pronounced for larger numbers of data points.
+They also note that this function is not concave, and so there could be multiple local optima.
+However, they suggest that for larger data sets, one local optimum will generally be orders of magnitude more probable than the others, and that this can generally be assumed to be the correct choice of hyperparameters.
+
+For all except the exponential function, we calculate the gradient $\frac{\partial K} {\partial \bftheta}$ and use a gradient-based descent method.
+
+Note that $K_y \in \R^{n \times n}$ must be recalculated at each step of the optimization.
+As the size of the data set grows, this becomes increasingly intractible.
+Therefore, we follow the suggestion of Russell \& Williams in their example (Ch 2), and calculate the hyperparameters for a smaller subset of the data.
+
+We are unsure whether there are any temporal trends within our data, and so we randomly select values from the training set.
+However, we expect that the temporal effects should not be too large.
+This should smear out any temporal dependance, leaving us with time-averaged values of the hyperparameters.
+
 \subsubsection*{Acknowledgments}

### 2014-05-03 code/optimize_hyperparams.m
Intermediate commit: Try optimizing script.
Still is problematic -- the solution is probability > 1!!

(Not sure why.)

@@ -3,36 +3,48 @@ function [theta, fval] = optimize_hyperparams(X,y,cov_type)
 addpath covar_hyperparams/.
 
 n_theta = 29;
-theta_0 = ones(n_theta, 1);
+%theta_0 = chi2rnd(3, n_theta, 1); % This distribution has it's mode at 1.
+theta_0 = abs(randn(n_theta, 1));
 
 Xt = X'; % Each column of X is a 27-param table entries
 %y is a column vector
 
 % Lower bound - bound all to be >= 0
 lb = zeros(n_theta,1);
+ub = 50 * ones(n_theta,1);
+%lb(n_theta) = 0;
+%ub(n_theta) = 1e-8;
 
 switch cov_type
 	case 1  % Squared exponential
 		options = optimoptions(@fmincon,...
 			'Display','iter', ...
 			'GradObj','on', ...
-			'Algorithm','interior-point'); %trust-region-reflective');
+            'MaxFunEvals',5000, ...
+            'MaxIter', 1500, ...
+			'Algorithm','interior-point');
 		objective = @(theta) se_cov_obj(Xt,y,theta);
 	case 2  % Matern (1/2) [exponential covariance]
 		options = optimoptions(@fmincon,...
 			'Display','iter', ...
+            'MaxFunEvals',5000, ...
+            'MaxIter', 1500, ...
 			'Algorithm','interior-point');  % or trust-region-reflective
 		objective = @(theta) exponential_cov_obj(Xt,y,theta);
 	case 3  % Matern (3/2)
 		options = optimoptions(@fmincon,...
 			'Display','iter', ...
 			'GradObj','on', ...
+            'MaxFunEvals',5000, ...
+            'MaxIter', 1500, ...
 			'Algorithm', 'interior-point'); %'trust-region-reflective');
 		objective = @(theta) matern_1p5_cov_obj(Xt,y,theta);
 	case 4  % Matern (5/2)
 		options = optimoptions(@fmincon,...
 			'Display','iter', ...
 			'GradObj','on', ...
+            'MaxFunEvals',5000, ...
+            'MaxIter', 1500, ...
 			'Algorithm', 'interior-point'); %'trust-region-reflective');
 		objective = @(theta) matern_2p5_cov_obj(Xt,y,theta);
 	otherwise

### 2014-05-03 code/run_optimization.m

@@ -1,4 +1,4 @@
-%function [theta, y_val] = run_optimization()
+function [theta, y_val] = run_optimization()
 
 %Call readProcessedData to get our data from the file into y and X
 [y, X] = readProcessedData();
@@ -17,29 +17,74 @@ end
 
 % Run the optimizations several times
 D = size(X, 2);
-%theta = zeros(D+2,4,4);
-%fval = zeros(4,4);
+theta = zeros(D+2,4,4);
+fval = zeros(4,4);
 
-for cov_type = 2:4
- 	[theta(:, cov_type, 1), fval(cov_type, 1)] = optimize_hyperparams( Scaled_X(1:100,:), Scaled_y(1:100), cov_type);
- 	[theta(:, cov_type, 2), fval(cov_type, 2)] = optimize_hyperparams( Scaled_X(101:200,:), Scaled_y(101:200), cov_type);
- 	randSet = randperm(nData, 100);
- 	[theta(:, cov_type, 3), fval(cov_type, 3)] = optimize_hyperparams( Scaled_X(randSet,:), Scaled_y(randSet), cov_type);
+randSet = randperm(N, 100);
 
-	% Also - check to see if this makes things better.  (Scale the sample separately.)
-	% This would mean that we would scale 1-10000, then rescale at each time step.
-	SampleSize = 100;
-	Scaled_X = zeros(SampleSize, D);
-	mean_y = mean(y(1:SampleSize));
-	stdev_y = std(y(1:SampleSize));
-	mean_X = mean(X(1:SampleSize, :),1);
-	stdev_X = std(X(1:SampleSize, :),0,1);
-	Scaled_y = y - mean_y;
-	for i = 1:N
-		Scaled_X(i,:) = (X(i,:) - mean_X) ./ stdev_X;
-	end
-	[theta(:, cov_type, 4), fval(cov_type, 4)] = optimize_hyperparams( Scaled_X(1:100,:), Scaled_y(1:100), cov_type);
+% Specially scaled 1-100
+SampleSize = 100;
+local_Scaled_X = zeros(SampleSize, D);
+local_mean_y = mean(y(1:SampleSize));
+local_stdev_y = std(y(1:SampleSize));
+local_mean_X = mean(X(1:SampleSize, :),1);
+local_stdev_X = std(X(1:SampleSize, :),0,1);
+local_Scaled_y = y - local_mean_y;
+for i = 1:N
+    local_Scaled_X(i,:) = (X(i,:) - local_mean_X) ./ local_stdev_X;
+end
 
[…]

### 2014-05-03 code/covar_hyperparams/exponential_cov_obj.m
Add debug method

@@ -1,7 +1,12 @@
 function obj = exponential_cov_obj( X, y, theta)
 % Negative Marginal likelihood of exponential covariance (objective function to be MINimized)
 
-K = K_mat(X, theta);
+if exist('debug.txt','file')
+	theta'
+	keyboard
+end
+
+K = K_mat(theta);
 try
 	Kinv = pinv(K);
 catch exception

### 2014-05-03 code/covar_hyperparams/matern_1p5_cov_obj.m

@@ -1,6 +1,11 @@
 function [obj, grad] = matern_1p5_cov_obj( X, y, theta)
 % Negative Marginal likelihood of Matern covariance (nu = 3/2) (objective function to be MINimized)
 
+if exist('debug.txt','file')
+	theta'
+	keyboard
+end
+
 K = K_mat(X, theta);
 try
 	Kinv = pinv(K);

### 2014-05-03 code/covar_hyperparams/matern_2p5_cov_obj.m

@@ -1,6 +1,11 @@
 function [obj, grad] = matern_2p5_cov_obj( X, y, theta)
 % Negative Marginal likelihood of Matern covariance (nu = 5/2) (objective function to be MINimized)
 
+if exist('debug.txt','file')
+	theta'
+	keyboard
+end
+
 K = K_mat(X, theta);
 try
 	Kinv = pinv(K);

### 2014-05-03 code/covar_hyperparams/se_cov_obj.m

@@ -1,6 +1,11 @@
 function [obj, grad] = se_cov_obj( X, y, theta)
 % Negative Marginal likelihood of squared exponential covariance (SE) (objective function to be MINimized)
 
+if exist('debug.txt','file')
+	theta'
+	keyboard
+end
+
 K = K_mat(X, theta);
 try
 	Kinv = pinv(K);
@@ -9,9 +14,11 @@ catch exception
 end
 
 [d, N] = size(X);
+
+%theta'
+%fprintf('term 1: %f, term 2: %f, term 3: %f\n', -0.5*y'*Kinv*y, -0.5*sum(log(svd(K))), -N/2*log(2*pi));
 % sum(log(svd(K))) is equivalent to log(det(K)), but less likely to blow up
 obj = - 0.5 * y' * Kinv * y - 0.5 * sum(log(svd(K))) - N/2 * log(2 * pi);
-%fprintf('term 1: %f, term 2: %f, term 3: %f\n', -0.5*y'*Kinv*y, -0.5*sum(log(svd(K))), -N/2*log(2*pi));
 grad = se_cov_grad(X, y, theta, Kinv);
 
 % Use negative of obj and grad, since we have to minimize

### 2014-05-03 code/run_optimization.m
Clean up run_optimization script

@@ -17,42 +17,26 @@ end
 
 % Run the optimizations several times
 D = size(X, 2);
-theta = zeros(D+2,4,4);
-fval = zeros(4,4);
+theta = zeros(D+2,4,3);
+fval = zeros(4,3);
 
 randSet = randperm(N, 100);
 
-% Specially scaled 1-100
-SampleSize = 100;
-local_Scaled_X = zeros(SampleSize, D);
-local_mean_y = mean(y(1:SampleSize));
-local_stdev_y = std(y(1:SampleSize));
-local_mean_X = mean(X(1:SampleSize, :),1);
-local_stdev_X = std(X(1:SampleSize, :),0,1);
-local_Scaled_y = y - local_mean_y;
-for i = 1:N
-    local_Scaled_X(i,:) = (X(i,:) - local_mean_X) ./ local_stdev_X;
-end
-
-for cov_type = 1:1
-    for data_set = 1:1
+for cov_type = 1:4
+    for data_set = 1:3
         switch data_set
             case 1 % Data between 1 and 100
                 set = 1:100;
                 X_set = Scaled_X(set,:);
                 y_set = Scaled_y(set);
-            case 2 % Data between 10001 and 10100
-                set = 10001:10100;
+            case 2 % Data between 8001 and 8100
+                set = 8001:8100;
                 X_set = Scaled_X(set,:);
                 y_set = Scaled_y(set);
             case 3 % Randomly selected data
                 set = randSet;
                 X_set = Scaled_X(set,:);
                 y_set = Scaled_y(set);
-            case 4 % Data between 1 and 100, normalized specifically for this case
-                set = 1:100;
-                X_set = local_Scaled_X(set, :);
-                y_set = local_Scaled_y;
         end
         local_optima = [];
         local_optimizers = [];
@@ -84,6 +68,7 @@ for cov_type = 1:1
         [m, i] = max(local_optima);
         theta(:, cov_type, data_set) = m;
         fval(:, cov_type, data_set) = local_optima(:,i);
+        fprintf('local optimum: %f', m);
     end
 end
 

### 2014-05-03 code/covar_hyperparams/exponential_cov_grad.m
Try not recalculating Xi-Xj over and over again
This will hopefully speed things up, at least a little bit.

Could actually save the 'u' values in grad for even greater savings.

@@ -1,10 +1,11 @@
-function grad = exponential_cov_grad(X, y, theta, K_inv)
+function grad = exponential_cov_grad(y, theta, K_inv)
 
 alpha = K_inv * y;
 term1 = (alpha * alpha' - K_inv);
 
 grad = zeros(numel(theta), 1);
-[d,N] = size(X);
+global X_prime;
+N = size(X_prime,1);
 
 for h = 1:numel(theta)
 	dK = zeros(N,N);
@@ -14,13 +15,13 @@ for h = 1:numel(theta)
 				if (i == j)
 					% So, add a tiny bit of noise to X_i.
 					% Not sure that this is legit, really.
-					x_temp = X(:,j)+1e-6.*randn(d,1);
-					dK(j,i) = dk_func_ll(X(:,i), x_temp, theta, h);
+					x_temp = 1e-6.*randn(d,1);
+					dK(j,i) = dk_func_ll(x_temp, theta, h);
 				else
-					dK(j,i) = dk_func_ll(X(:,i), X(:,j), theta, h);
+					dK(j,i) = dk_func_ll(reshape(X_prime(i,j,:),D,1), theta, h);
 				end
 			elseif h == d + 1
-				dK(j,i) = dk_func_sigma_f(X(:,i), X(:,j), theta);
+				dK(j,i) = dk_func_sigma_f(reshape(X_prime(i,j,:),D,1), theta);
 			elseif h == d + 2
 				dK(j,i) = dk_func_sigma_n(theta, (i==j) );
 			end
@@ -33,19 +34,16 @@ end
 
 
 
-function out = dk_func_ll(x1, x2, theta, j)
+function out = dk_func_ll(d, theta, j)
 % Derivative of Matern covariance function with respect to l[j] hyperparameter
-%    x1 - first coordinate
-%    x2 - second coordinate
+%    d - difference in coordinates
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
 %       element d+1: sigma_f
 %       element d+2: sigma_n
 %    j - hyperparameter
 
-D = numel(x1);
-
-d = x1 - x2;
+D = numel(d);
 u = sqrt(d' * diag(1 ./ (theta(1:D) .^2)) * d);
 
 % d/dl sigma_f * exp(-u) = sigma_f * exp(-u) * -(1/(2 * sqrt(sum(d^2/ll^2)))) * (-2 d_j^2 / l_j^3)
[…]

### 2014-05-03 code/covar_hyperparams/exponential_cov_obj.m

@@ -1,4 +1,4 @@
-function obj = exponential_cov_obj( X, y, theta)
+function obj = exponential_cov_obj(theta, y)
 % Negative Marginal likelihood of exponential covariance (objective function to be MINimized)
 
 if exist('debug.txt','file')
@@ -13,11 +13,11 @@ catch exception
 	Kinv = pinv(K + 1e-8 * randn(size(K)));
 end
 
-[d, N] = size(X);
+[N, ~] = size(K);
 % sum(log(svd(K))) is equivalent to log(det(K)), but less likely to blow up
 obj = - 0.5 * y' * Kinv * y - 0.5 * sum(log(svd(K))) - N/2 * log(2 * pi);
 
-%grad = exponential_cov_grad(X, y, theta, Kinv);
+%grad = exponential_cov_grad(y, theta, Kinv);
 
 % Use negative of obj and grad, since we have to minimize
 obj = - obj;
@@ -26,29 +26,29 @@ obj = - obj;
 
 
 
-function out = K_mat(X, theta)
+function out = K_mat(theta)
 % Matern covariance function matrix (objective function)
-%    X - data matrix
 %    theta - the struct of coordinates that can be messed with.
 %       Contains ll, a vector of scaling params, and sigma_f and sigma_n
 
-[d, N] = size(X);
+global X_prime;
+N = size(X_prime, 1);
+D = size(X_prime, 3);
 out = zeros(N);
 for i = 1:N
-	out(i,i) =  k_func(X(:,i), X(:,i), theta, true);
+	out(i,i) =  k_func(reshape(X_prime(i,i,:),D,1), theta, true);
 	for j = i+1:N
-		out(i,j) = k_func(X(:,i), X(:,j), theta, false);
+		out(i,j) = k_func(reshape(X_prime(i,j,:),D,1), theta, false);
 		out(j,i) = out(i,j);
-	end
+    end
 end
 
 
 
 
-function k = k_func(x1, x2, theta, on_diagonal)
+function k = k_func(d, theta, on_diagonal)
 % Matern covariance function
-%    x1 - first coordinate
-%    x2 - second coordinate
+%    d - x1 - x2
 %    theta - the array of coordinates that can be changed.
 %       elements 1:d :  l, a vector of scaling params
 %       element d+1: sigma_f
[…]

### 2014-05-03 code/optimize_hyperparams.m

@@ -11,9 +11,22 @@ Xt = X'; % Each column of X is a 27-param table entries
 
 % Lower bound - bound all to be >= 0
 lb = zeros(n_theta,1);
-ub = 50 * ones(n_theta,1);
+%ub = 50 * ones(n_theta,1);
 %lb(n_theta) = 0;
 %ub(n_theta) = 1e-8;
+
+D = size(Xt, 2);
+
+global X_prime;
+X_prime = zeros(D,D,27);
+
+for (i = 1:D)
+	for j = i:D
+		X_prime(i,j,:) = Xt(:,i) - Xt(:,j);
+		X_prime(j,i,:) = X_prime(i,j,:);
+	end
+end
+
 
 switch cov_type
 	case 1  % Squared exponential
@@ -30,7 +43,7 @@ switch cov_type
             'MaxFunEvals',5000, ...
             'MaxIter', 1500, ...
 			'Algorithm','interior-point');  % or trust-region-reflective
-		objective = @(theta) exponential_cov_obj(Xt,y,theta);
+		objective = @(theta) exponential_cov_obj(theta,y);
 	case 3  % Matern (3/2)
 		options = optimoptions(@fmincon,...
 			'Display','iter', ...

### 2014-05-07 code/run_optimization.m
Fix optimization script

@@ -1,4 +1,4 @@
-function [theta, y_val] = run_optimization()
+function [theta, y_val] = run_optimization(type, set, amt, num)
 
 %Call readProcessedData to get our data from the file into y and X
 [y, X] = readProcessedData();
@@ -20,17 +20,17 @@ D = size(X, 2);
 theta = zeros(D+2,4,3);
 fval = zeros(4,3);
 
-randSet = randperm(N, 100);
+randSet = randperm(10000, amt);
 
-for cov_type = 1:4
-    for data_set = 1:3
+for cov_type = type:type
+    for data_set = set:set
         switch data_set
             case 1 % Data between 1 and 100
-                set = 1:100;
+                set = 1:amt;
                 X_set = Scaled_X(set,:);
                 y_set = Scaled_y(set);
             case 2 % Data between 8001 and 8100
-                set = 8001:8100;
+                set = 8001:8000+amt;
                 X_set = Scaled_X(set,:);
                 y_set = Scaled_y(set);
             case 3 % Randomly selected data
@@ -57,20 +57,25 @@ for cov_type = 1:4
             local_optimizers(:,iter) = t;
             novel = true;
             for i = 1:iter-1
-                if (norm(t - local_optimizers(:,i)) < 1e-2)
+                if (norm(t - local_optimizers(:,i)) < 1e-1)
                     novel = false;
                 end
             end
             if (novel == true)
                 lastNovel = iter;
             end
+	    fprintf('current optimum: %f', f);
+	    fprintf('current optimzer:');
+	    t'
+	    save(['optimized_t', num2str(num)], 't')
+	    save(['optimized_f', num2str(num)], 'f')
         end
         [m, i] = max(local_optima);
-        theta(:, cov_type, data_set) = m;
-        fval(:, cov_type, data_set) = local_optima(:,i);
+        theta(:, cov_type, data_set) = local_optimizers(:,i);
+        fval(:, cov_type, data_set) = local_optima(i);
         fprintf('local optimum: %f', m);
     end
 end
 
-save('optimized_theta', 'theta')
-save('optimized_fval', 'fval')
+save(['optimized_theta', num2str(num)], 'theta')
+save(['optimized_fval', num2str(num)], 'fval')

### 2014-05-07 code/covar_hyperparams/exponential_cov_obj.m
Optimizations for covariance

@@ -6,7 +6,7 @@ if exist('debug.txt','file')
 	keyboard
 end
 
-K = K_mat(theta);
+K = K_mat(X, theta);
 try
 	Kinv = pinv(K);
 catch exception

### 2014-05-07 code/Gaussian_Processes_Regression.m
Uncommented strategy 0 and renamed predictedy_values.txt

@@ -8,9 +8,9 @@ stdev_y = std(y);
 
 %Strategy 1: Keep those rows for which mean_y - stdev_y <= y <= mean_y +
 %stdev_y
-y_inliers = (mean_y - stdev_y) <= y & y <= (mean_y + stdev_y);
-Pruned_X = X(y_inliers, :);
-Pruned_y = y(y_inliers);
+% y_inliers = (mean_y - stdev_y) <= y & y <= (mean_y + stdev_y);
+% Pruned_X = X(y_inliers, :);
+% Pruned_y = y(y_inliers);
 
 
 %Strategy 2: Keep those rows for which y <= mean_y - stdev_y and y >= 
@@ -20,8 +20,8 @@ Pruned_y = y(y_inliers);
 % Pruned_y = y(y_outliers);
 
 %Strategy 3: No strategies
-% Pruned_X = X;
-% Pruned_y = y;
+Pruned_X = X;
+Pruned_y = y;
 
 %Calculate the mean and std_dev of the pruned X and y
 mean_Pruned_X = mean(Pruned_X,1);
@@ -156,6 +156,7 @@ while (ctr < size(Design_X,2))
 end %end of while
 
 fileName = 'Predicted_yValues.txt';
+fileName = strcat(fileName, '_', num2str(windowSize),'_', num2str(cov_type), '.txt');
 writeFile(Predicted_yValues, fileName);
 end %end of function
 

### 2014-05-07 code/Gaussian_Processes_Regression.m
Uses the optimized hyperparameters

@@ -43,8 +43,12 @@ X_Window = Design_X(:,1:windowSize); y_Window = Design_y(1:windowSize);
 ctr = windowSize;
 
 %Select the hyper-parameters
-sigma_f = 1.0; sigma_n = 0.1;
-l = ones(size(X_Window,1),1)';   %All characteristic length scales are initialized to 1
+hyper_params_mat = load('theta_oh_250.mat');
+hyper_params = hyper_params_mat.theta(:);
+%sigma_f = 1.0; sigma_n = 0.1;
+%l = ones(size(X_Window,1),1)';   %All characteristic length scales are initialized to 1
+l = hyper_params(1:27,1)';
+sigma_f = hyper_params(28); sigma_n = hyper_params(29);
 
 %Check to see whether K is stored. If not, build the matrix from scratch
 num_examples = size(X_Window,2);
@@ -86,7 +90,6 @@ end
 
 
 %Train the hyperparameters of the initial K matrix
-
 
 Predicted_yValues = [];
 while (ctr < size(Design_X,2))        

### 2014-05-07 code/Gaussian_Processes_Regression.m
Removed .txt from Predicted_yValues

@@ -158,7 +158,7 @@ while (ctr < size(Design_X,2))
 
 end %end of while
 
-fileName = 'Predicted_yValues.txt';
+fileName = 'Predicted_yValues';
 fileName = strcat(fileName, '_', num2str(windowSize),'_', num2str(cov_type), '.txt');
 writeFile(Predicted_yValues, fileName);
 end %end of function

### 2014-05-07 code/Gaussian_Processes_Regression.m
Changed the while loop condition

@@ -92,7 +92,8 @@ end
 %Train the hyperparameters of the initial K matrix
 
 Predicted_yValues = [];
-while (ctr < size(Design_X,2))        
+%while (ctr < size(Design_X,2))        
+while (ctr < (windowSize + 500))        
     
     %Build the delta Gram matrix
     if (ctr ~= windowSize)

### 2014-05-07 code/Gaussian_Processes_Regression.m
Updated GPR to have strategy type

@@ -1,4 +1,4 @@
-function Gaussian_Processes_Regression(windowSize, cov_type)
+function Gaussian_Processes_Regression(windowSize, cov_type, strategy_type)
 
 %Call readProcessedData to get our data from the file into y and X 
 [y, X] = readProcessedData();
@@ -6,22 +6,19 @@ function Gaussian_Processes_Regression(windowSize, cov_type)
 mean_y = mean(y);
 stdev_y = std(y);
 
-%Strategy 1: Keep those rows for which mean_y - stdev_y <= y <= mean_y +
-%stdev_y
-% y_inliers = (mean_y - stdev_y) <= y & y <= (mean_y + stdev_y);
-% Pruned_X = X(y_inliers, :);
-% Pruned_y = y(y_inliers);
-
-
-%Strategy 2: Keep those rows for which y <= mean_y - stdev_y and y >= 
-%mean_y + stdev_y
-% y_outliers = y <= (mean_y - stdev_y) & y >= (mean_y + stdev_y);
-% Pruned_X = X(y_outliers, :);
-% Pruned_y = y(y_outliers);
-
-%Strategy 3: No strategies
-Pruned_X = X;
-Pruned_y = y;
+switch cov_type
+	case 1 %Strategy 1: Keep those rows for which mean_y - stdev_y <= y <= mean_y + stdev_y
+		y_inliers = (mean_y - stdev_y) <= y & y <= (mean_y + stdev_y);
+		Pruned_X = X(y_inliers, :);
+		Pruned_y = y(y_inliers);
+	case 2 %Strategy 2: Keep those rows for which y <= mean_y - stdev_y and y >=  mean_y + stdev_y
+		y_outliers = y <= (mean_y - stdev_y) & y >= (mean_y + stdev_y);
+		Pruned_X = X(y_outliers, :);
+		Pruned_y = y(y_outliers);
+	case 0 % aka Strategy 3:  No strategies
+		Pruned_X = X;
+		Pruned_y = y;
+end
 
 %Calculate the mean and std_dev of the pruned X and y
 mean_Pruned_X = mean(Pruned_X,1);
@@ -43,7 +40,13 @@ X_Window = Design_X(:,1:windowSize); y_Window = Design_y(1:windowSize);
 ctr = windowSize;
 
 %Select the hyper-parameters
-hyper_params_mat = load('theta_oh_250.mat');
+hyper_params = hyper_params_mat.theta(:);
+switch cov_type
+case 1
[…]

### 2014-05-07 code/Gaussian_Processes_Regression.m
Fix GPR things

@@ -6,7 +6,10 @@ function Gaussian_Processes_Regression(windowSize, cov_type, strategy_type)
 mean_y = mean(y);
 stdev_y = std(y);
 
-switch cov_type
+switch strategy_type
+	case 0 % aka Strategy 3:  No strategies
+		Pruned_X = X;
+		Pruned_y = y;
 	case 1 %Strategy 1: Keep those rows for which mean_y - stdev_y <= y <= mean_y + stdev_y
 		y_inliers = (mean_y - stdev_y) <= y & y <= (mean_y + stdev_y);
 		Pruned_X = X(y_inliers, :);
@@ -15,9 +18,8 @@ switch cov_type
 		y_outliers = y <= (mean_y - stdev_y) & y >= (mean_y + stdev_y);
 		Pruned_X = X(y_outliers, :);
 		Pruned_y = y(y_outliers);
-	case 0 % aka Strategy 3:  No strategies
-		Pruned_X = X;
-		Pruned_y = y;
+	otherwise
+		fprintf('error: invalid strategy');
 end
 
 %Calculate the mean and std_dev of the pruned X and y
@@ -40,12 +42,13 @@ X_Window = Design_X(:,1:windowSize); y_Window = Design_y(1:windowSize);
 ctr = windowSize;
 
 %Select the hyper-parameters
-hyper_params = hyper_params_mat.theta(:);
 switch cov_type
 case 1
 	hyper_params_mat = load('theta_se_500.mat');
 case 2
 	hyper_params_mat = load('theta_oh_500.mat');
+otherwise
+	fprintf('error: invalid cov type');
 end
 hyper_params = hyper_params_mat.theta(:);
 %sigma_f = 1.0; sigma_n = 0.1;
@@ -57,7 +60,7 @@ sigma_f = hyper_params(28); sigma_n = hyper_params(29);
 num_examples = size(X_Window,2);
 K = zeros(num_examples);
 fileName = 'K_Matrix';
-fileName = strcat(fileName, '_', num2str(windowSize),'_', num2str(cov_type), '.txt');
+fileName = strcat(fileName, '_', num2str(windowSize),'_', num2str(cov_type), 'trained_with_500vals__', num2str(strategy_type),'.txt');
 path = '../data/';
 if exist([path fileName], 'file')
    K = dlmread([path fileName],'\t');

### 2014-05-07 code/Gaussian_Processes_Regression.m
Fix strategy 2

@@ -14,8 +14,8 @@ switch strategy_type
 		y_inliers = (mean_y - stdev_y) <= y & y <= (mean_y + stdev_y);
 		Pruned_X = X(y_inliers, :);
 		Pruned_y = y(y_inliers);
-	case 2 %Strategy 2: Keep those rows for which y <= mean_y - stdev_y and y >=  mean_y + stdev_y
-		y_outliers = y <= (mean_y - stdev_y) & y >= (mean_y + stdev_y);
+	case 2 %Strategy 2: Keep those rows for which y <= mean_y - stdev_y OR y >=  mean_y + stdev_y
+		y_outliers = y <= (mean_y - stdev_y) | y >= (mean_y + stdev_y);
 		Pruned_X = X(y_outliers, :);
 		Pruned_y = y(y_outliers);
 	otherwise

### 2014-05-07 code/Gaussian_Processes_Regression.m
Try this to fix K

@@ -148,7 +148,12 @@ while (ctr < (windowSize + 500))
                 error('Wrong covariance function type. Please try again!');                    
         end
     end
-    L = chol(K, 'lower');
+    try
+        L = chol(K, 'lower');
+    catch exception
+	save('K','K');
+        L = inv(K + 1e-5 * abs(normrnd(0, 1, size(K))));
+    end
     alpha = L'\(L\y_Window');
     %We need to rescale the predicted y by adding back the mean
     predicted_y = (k_star' * alpha) + mean_Pruned_y;      % Also, our predicted change in price is the mean of the gaussian posterior

### 2014-05-09 code/graphs.m
graphs, beginning

@@ -1,0 +1,59 @@
+function graphs()
+% Reads in data from the file and stores it as a matrix.
+addpath '../data/';
+folders = dir('../data');
+for i = 1 : numel(folders)
+    if (folders(i).isdir == 1 && ~strcmp(folders(i).name,'.') && ~strcmp(folders(i).name,'..'))
+        p = ['../data/', folders(i).name, '/']; % path
+        if exist([p, 'processed_data.txt'], 'file')
+        
+            [y, ~] = readProcessedData(p);
+
+            files = dir(p);
+            split_folder_name = regexp(folders(i).name,'_','split');       
+            n_train_data = str2num(split_folder_name{2});
+            y_training = y(1:n_train_data);
+            y_result = y(n_train_data+1:n_train_data+500);
+            next = 0;
+            for j = 1 : numel(files)
+                if files(j).isdir == 0 && (strcmp('Predicted_yValues',files(j).name(1:17)) == 1)
+                    next = next + 1;
+                    yPredicted(next,:) = readYPred([p, files(j).name]);
+                end
+            end
+            figure;
+            plot (1:n_train_data, y_training,'k.');
+            hold
+            title('Plot for training with 20000 samples')
+            xlabel('sample #')
+            ylabel('Change in price')
+            legend('Training values','Test values')
+            plot (n_train_data+1:n_train_data+500, y_result,'bo');
+            for j = 1 : next
+                plot (n_train_data+1:n_train_data+500, yPredicted(j,:),'c*');
+            end
+            saveas(gcf, 'fig_i.png');
+        end
+        
+        
+    end
+end
+
+
+
+fileName = 'processed_data.txt';
+
+fileID = fopen(fileName,'r');
+formatSpec = '%f';
+nCols = 33-6+1;
+for i = 2:nCols
+    formatSpec = [formatSpec ' %f'];
+end
+
+data = fscanf(fileID, formatSpec, [nCols Inf]);
+data = data';
+
+fclose(fileID);
+
+y = data(:,1);
+X = data(:,2:nCols);

### 2014-05-09 code/readYPred.m

@@ -1,0 +1,12 @@
+function [y] = readYPred(fileName)
+% Reads in data from the file and stores it as a vector.
+
+fileID = fopen(fileName,'r');
+formatSpec = '%f';
+nCols = 1;
+
+y = fscanf(fileID, formatSpec, [nCols Inf]);
+y = y';
+
+fclose(fileID);
+

### 2014-05-09 code/graphs.m
Graphs, part 2

@@ -2,58 +2,236 @@ function graphs()
 % Reads in data from the file and stores it as a matrix.
 addpath '../data/';
 folders = dir('../data');
-for i = 1 : numel(folders)
-    if (folders(i).isdir == 1 && ~strcmp(folders(i).name,'.') && ~strcmp(folders(i).name,'..'))
-        p = ['../data/', folders(i).name, '/']; % path
-        if exist([p, 'processed_data.txt'], 'file')
-        
-            [y, ~] = readProcessedData(p);
+close all;
 
-            files = dir(p);
-            split_folder_name = regexp(folders(i).name,'_','split');       
-            n_train_data = str2num(split_folder_name{2});
-            y_training = y(1:n_train_data);
-            y_result = y(n_train_data+1:n_train_data+500);
-            next = 0;
-            for j = 1 : numel(files)
-                if files(j).isdir == 0 && (strcmp('Predicted_yValues',files(j).name(1:17)) == 1)
-                    next = next + 1;
-                    yPredicted(next,:) = readYPred([p, files(j).name]);
-                end
-            end
-            figure;
-            plot (1:n_train_data, y_training,'k.');
-            hold
-            title('Plot for training with 20000 samples')
-            xlabel('sample #')
-            ylabel('Change in price')
-            legend('Training values','Test values')
-            plot (n_train_data+1:n_train_data+500, y_result,'bo');
-            for j = 1 : next
-                plot (n_train_data+1:n_train_data+500, yPredicted(j,:),'c*');
-            end
-            saveas(gcf, 'fig_i.png');
-        end
-        
-        
+%        deleted:    data/Predicted_yValues_10000_2trained_with_500vals__0.txt
+%         deleted:    data/Predicted_yValues_15000_1trained_with_500vals__0.txt
+%         deleted:    data/Predicted_yValues_15000_2_trained_with_500vals__0__msecs800.txt
+%         deleted:    data/Predicted_yValues_15000_2trained_with_500vals__0.txt
[…]

### 2014-05-09 code/graphs.m
Update graphs

@@ -181,7 +181,34 @@ saveas(gcf, ['fig_', num2str(plotNum), '.png']);
 
 %LR, NN, PCA Reg vs GPR. Keep window size fixed at 20000, timesecs_diff at 500,
 %strategy_type fixed at 0 and Ornstein-Uhlenbeck covariance function 
+barnames = ['Predicted_yValues_10000_2trained_with_500vals__0.txt'; ...
+'Predicted_yValues_15000_2trained_with_500vals__0_1000msecs_diff.txt'; ...
+'Predicted_yValues_15000_1trained_with_500vals__0.txt'; ...
+'Predicted_yValues_15000_2trained_with_500vals__0.txt'; ...
+'Predicted_yValues_15000_2trained_with_500vals__0_500msecs_diff.txt'; ... 
+'Predicted_yValues_15000_2trained_with_500vals__1.txt'; ...
+'Predicted_yValues_15000_2_trained_with_500vals__0__msecs800.txt'; ...
+'Predicted_yValues_20000_1trained_with_500vals__0.txt'; ...
+'Predicted_yValues_20000_3trained_with_500vals__0.txt'; ...
+'Predicted_yValues_20000_4_trained_with_500vals_rand_0.txt'; ...
+'Predicted_yValues_3000_1trained_with_500vals__2.txt'; ...
+'Predicted_yValues_3000_2trained_with_500vals__2.txt'];
 
+bars = [11.08, -10.96, 3.80; ...
+    3.86, 4.14, 4.02; ...
+    3.86, 4.14, 4.02; ...
+    3.86, 4.14, 4.02; ...
+    3.86, 4.14, 4.02; ...
+    42.88, -57.00, -74.32; ...
+    3.86, 4.14, 4.02; ...
+    12.87, -8.92, -16.53; ...
+    12.87, -8.92, -16.53; ...
+    12.87, -8.92, -16.53; ...
+    28.39, -61.22, -76.70; ...
+    28.39, -61.22, -76.70];
+legend('Our model', 'LM', 'PCA\_reg');
+
+bar(1:12, bars)
 
 
 clear;

### 2014-05-10 paper/paper.tex

--- 2014-04-23 paper/paper.tex
@@
 \section{Introduction}
-Financial data is fun.
+In stock markets, investors need long-term forecasting techniques to choose the right time to buy/sell stocks to maximize their profits or to minimize their loss.
+The majority of existing stock market forecasting techniques rely on machine learning tools and the accuracy of these tools is still a significant challenge for market traders considering the several external factors which affect the stock market.
+In this project we have tried to forecast change in price of a single stock using the Gaussian Processes (GP) model.
+We have also investigated the use of Gaussian processes with different covariance functions to model real world temporal patterns.
+We
 
@@
 We sort the data set in increasing time.
-For each line, we define $\delta$, the difference in stock price after 2 minutes, to be the current stock price subtracted from the stock price that is associated with the datapoint that is closest to 2 minutes in the future.
-(Note that our datapoints are roughly 50 milliseconds apart, so we can expect that the variation in distance between sets of data points will be much less than modeling error.)
+For each line, we define $\delta$, the difference in stock price after \textbf{t} milliseconds, to be the current stock price subtracted from the stock price that is associated with the datapoint that is closest to \textbf{t} milliseconds in the future.
+(Note that our datapoints are roughly 500 milliseconds apart, so we can expect that the variation in distance between sets of data points will be much less than modeling error.)
 We will attempt to build a model which calculates $\delta_t$ when given $P_t$, the values of the 27 parameters at time $t$.
@@
 % Note to self: this is done in strategy
+
+% TODO:  Insert a paragraph about normalization of data.
+
+\section{Algorithm}
+
+Our algorithm could be divided into 4 major parts - preprocessing the data, building the covariance matrix, tuning the hyperparameters, learning the GP model using a windowing technique and finally predicting the change in stock price for the next time instant using the learned GP model.
+
+\subsection{Pre-processing}
+
+We experiment with 3 different strategies to build our training set from the available dataset.
+In the first strategy we prune out all data points that were more than one standard deviation away from the sample mean $\bf\bar{y}$.
+In the second strategy we prune out all data points that were within one standard deviation from the sample mean $\bf\bar{y}$.
+And for the third strategy, we keep all data points.
+After applying any one of the above strategies to the data, we linearly scale the inputs to have zero mean and unit variance on the training set.
+We also centered the corresponding target outputs so as to have zero mean on the training set.
+
+\subsection{Building the Covariance Matrix}
+
+We initially build the covariance matrix - $K$ using the first $T$ samples, i.e using the samples starting at $t = 0$ till $t = T$.
+While building $K$, we have experimented with a few interesting covariance functions as described below:
+
+\subsubsection{Covariance Functions}
+We consider 4 different covariance functions, all of which can be parameterized in terms of the hyperparameters $\bftheta = \{\bfl, \sigma_n^2 \sigma_f^2\}$:
+Let
+\begin{equation}
+	u = \left( \left( \bfx_p - \bfx_q \right)^\top \diag(\bfl)^{-2} \left( \bfx_p - \bfx_q \right) \right)^{1/2}
+\end{equation}
+
+Then, the covariance functions are:
+\begin{align}
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- \frac{1}{2} u^2} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Squared Exponential}   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Exponential}   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{3} u \right) \exp{- sqrt{3}  u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{3}{2}$) }   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{5} u \right) + 5 u^2) \exp{- \sqrt{5} u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{5}{2}$) }
+\end{align}
+
+$\sigma_n^2$ and $\sigma_f^2$ are the noise and signal level parameters, respectively, and $l_1, \dots, l_D$ are characteristic length-scales.
+The covariance matrix $K_y$ is given by $[K_y]_{pq} = k(\bfx_p, \bfx_q)$.
+We can think of these length-scales as a measure of the irrelevance of the associated feature; a large value of $l_i$ means that the $i$-th parameter of the feature vector has very little effect on the resultant covariance matrix.
+This is a type of automatic relevance determination (ARD).
+
+\subsection{Tuning the hyperparameters}
+We tuned the hyperparameters by maximizing the log marginal likelihood, 
+\begin{equation}
+	\log p(\bfy | X, \bftheta) = - \frac{1}{2} \bfy^\top K_y^{-1} \bfy - \frac{1}{2} \log \| K_y \| - \frac{n}{2} \log (2 \pi)
+\end{equation}
+
+Russell and Williams state that the maximum of this quantity is generally more pronounced for larger numbers of data points.
+They also note that this function is not concave, and so there could be multiple local optima.
+However, they suggest that for larger data sets, one local optimum will generally be orders of magnitude more probable than the others, and that this can generally be assumed to be the correct choice of hyperparameters.
+
+For all except the exponential function, we calculate the gradient $\frac{\partial K} {\partial \bftheta}$ and use a gradient-based descent method.
+
+Note that $K_y \in \R^{n \times n}$ must be recalculated at each step of the optimization.
+As the size of the data set grows, this becomes increasingly intractable.
+Therefore, we follow the suggestion of Russell \& Williams in their example (Ch 2), and optimize the hyperparameters based on a smaller subset of the data.
+We assume that there are temporal trends within the data, and thus calculate the marginal likelihood using the first 500 data points from the training set which we then feed as input to our optimizer.
+After we obtain the optimized set of hyperparameters, the same set is employed in all our experiments.
+
+\subsection{Learning the GP Model}
+
+We have applied a windowing technique to learn a GP on the data.
+We use the covariance matrix built using data from $t = 0$ till $t = \mathcal{T}$, to predict the change in stock price at time instant $t = \mathcal{T}+1$.
+Once the prediction is made, we use the input vector at $t = \mathcal{T}+1$ and the \textbf{\textit{predicted}} change in price at that instant as a new data point and compute the covariance matrix again for the window from $t = 1$ till $t = \mathcal{T}+1$.
+We use this newly computed covariance matrix to predict the change in stock price at time instant $t=\mathcal{T}+2$.
+This process is repeated till time instant $t = \mathcal{T}+T$, where $T$ denotes how far ahead in future we want to keep predicting the change in stock prices.
+Since re-computing the covariance matrix for every window can be pretty time consuming, we compute only the last column / row of the covariance matrix as it corresponds to the time instant $t = \mathcal{T}+1$.
+In addition, we drop the row and column that correspond to the oldest time instant of the previous window, from the covariance matrix.
+This keeps the size of our covariance matrix constant (therefore, keeping the amount of computation required for inverting the covariance matrix constant across every window) while allowing us to re-use the other columns of the covariance matrix that correspond to the time instants common to the previous window and the current window.
+
+\subsection{Prediction}
+
+We predict the change in stock price at the time instant immediately following our current time window, to be the \textbf{\textit{mean}} of the below conditional distribution:
+
+\begin{equation}
+ \boldsymbol{f_*} | \boldsymbol{X}, \boldsymbol{y}, \boldsymbol{X_*} \sim \mathcal{N}\left(\bar{\boldsymbol{f_*}}, \mbox{cov}(\boldsymbol{f_*})\right)
+\end{equation}
+where,
+\begin{align}
+\bar{\boldsymbol{f_*}} &= \boldsymbol{K}\left(\boldsymbol{X_*}, \boldsymbol{X}\right) \left[ \boldsymbol{K}\left(\boldsymbol{X}, \boldsymbol{X}\right) + \sigma_n^2\boldsymbol{I}\right]^{-1} \boldsymbol{y} \\
+\mbox{cov}(\boldsymbol{f_*}) &= \boldsymbol{K}\left(\boldsymbol{X_*}, \boldsymbol{X_*}\right) - \boldsymbol{K}\left(\boldsymbol{X_*}, \boldsymbol{X}\right) \left[ \boldsymbol{K}\left(\boldsymbol{X}, \boldsymbol{X}\right) + \sigma_n^2\boldsymbol{I}\right]^{-1} \boldsymbol{K}\left(\boldsymbol{X}, \boldsymbol{X_*}\right)
+\end{align}
+
+Finally we scale the predicted value back by adding the sample mean - $\bf\bar{y}$, of the target vector to our prediction.
+
+\section{Conclusion}
+
+The results from our experiments indicate that GP models could learn the correlations hidden among financial data to a very high degree when compared to other traditional ML algorithms like linear regression, PCA regression and neural networks.
+But the biggest drawback of GP models as implemented in our application is the computation time that is required for the GP modeling.
+Regression based on GP models involve a huge amount of matrix inversions, the complexity of which is to the order of $\mathcal{O}(n^3)$.
+These computational needs, restricted the amount of training data that could be used, to a maximum of only 20000 training samples, even though we had access to around 2GB of stock price data that we would have liked to use during training.
+Several authors have suggested various sparse approximation approaches towards calculating the covariance matrix in which they try to retain most of the information contained in the full dataset while reducing the size of the covariance matrix so as to facilitate a less demanding implementation of the GP model from the computational point of view.
 

### 2014-05-10 paper/paper.tex

@@
 
-\title{Formatting Instructions for NIPS 2014}
+\title{Applying Gaussian Process Regression to Predict Changes in Stock Price}
 
@@
 In this project we have tried to forecast change in price of a single stock using the Gaussian Processes (GP) model.
-We have also investigated the use of Gaussian processes with different covariance functions to model real world temporal patterns.
-We
 
@@
 We sort the data set in increasing time.
-For each line, we define $\delta$, the difference in stock price after \textbf{t} milliseconds, to be the current stock price subtracted from the stock price that is associated with the datapoint that is closest to \textbf{t} milliseconds in the future.
-(Note that our datapoints are roughly 500 milliseconds apart, so we can expect that the variation in distance between sets of data points will be much less than modeling error.)
+For each line, we define $\delta$, the difference in stock price after \textbf{t} milliseconds, to be the current stock price subtracted from the stock price that is closest to \textbf{t} milliseconds in the future.
+%(Note that our datapoints are close enough together that the difference in time between sets of data points will be much less than modeling error.)
 We will attempt to build a model which calculates $\delta_t$ when given $P_t$, the values of the 27 parameters at time $t$.
@@
 
-Our algorithm could be divided into 4 major parts - preprocessing the data, building the covariance matrix, tuning the hyperparameters, learning the GP model using a windowing technique and finally predicting the change in stock price for the next time instant using the learned GP model.
+Our algorithm could be divided into 4 major parts - preprocessing the data, building the covariance matrix, tuning the hyperparameters, learning the GP model using a windowing technique and finally predicting the change in stock price for the next time instance using the learned GP model.
 
@@
 
-We initially build the covariance matrix - $K$ using the first $T$ samples, i.e using the samples starting at $t = 0$ till $t = T$.
+We initially build the covariance matrix $K$ using the first $\mathcal{T}$ samples, i.e using the samples from $t = 0$ to $t = \mathcal{T}$.
 While building $K$, we have experimented with a few interesting covariance functions as described below:
@@
 \subsubsection{Covariance Functions}
-We consider 4 different covariance functions, all of which can be parameterized in terms of the hyperparameters $\bftheta = \{\bfl, \sigma_n^2 \sigma_f^2\}$:
+We consider 4 different covariance functions, all of which can be parameterized in terms of the hyperparameters $\bftheta = \{\bfl, \sigma_n^2 \sigma_f^2\}$, where $\bfl$ is a length 27 vector.
 Let
@@
 	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Exponential}   \\
-	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{3} u \right) \exp{- sqrt{3}  u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{3}{2}$) }   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{3} u \right) \exp{- \sqrt{3}  u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{3}{2}$) }   \\
 	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{5} u \right) + 5 u^2) \exp{- \sqrt{5} u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{5}{2}$) }
@@
 \subsection{Tuning the hyperparameters}
-We tuned the hyperparameters by maximizing the log marginal likelihood, 
+For the covariance matrices above, we tuned the hyperparameters by maximizing the log marginal likelihood, 
 \begin{equation}
@@
 
-Russell and Williams state that the maximum of this quantity is generally more pronounced for larger numbers of data points.
+Rasmussen and Williams state that the maximum of this quantity is generally more pronounced for larger numbers of data points.
 They also note that this function is not concave, and so there could be multiple local optima.
@@
 For all except the exponential function, we calculate the gradient $\frac{\partial K} {\partial \bftheta}$ and use a gradient-based descent method.
-
 Note that $K_y \in \R^{n \times n}$ must be recalculated at each step of the optimization.
 As the size of the data set grows, this becomes increasingly intractable.
-Therefore, we follow the suggestion of Russell \& Williams in their example (Ch 2), and optimize the hyperparameters based on a smaller subset of the data.
-We assume that there are temporal trends within the data, and thus calculate the marginal likelihood using the first 500 data points from the training set which we then feed as input to our optimizer.
+Therefore, we follow the suggestion of Rasmussen \& Williams in their example (Ch 2), and optimize the hyperparameters based on a smaller subset of the data.
+
+We calculated the marginal likelihood using the first 500 data points from the training set, and also using 500 random data points from the training set.
+We found that the K matrix was ill-conditioned when we used 500 random data points, so we did not use these values in our experiments.
+Ideally, we would have used a larger dataset and more random restarts, but we were unable to do so due to time constraints.
+We believe this is one of the major sources of error in our project.
 After we obtain the optimized set of hyperparameters, the same set is employed in all our experiments.
+
+We observed that the resultant $\bfl$ parameters had very large values in the majority of their coefficients and smaller values for a limited subset of components.
+This means that a limited number of components have a larger correlation with the target value.
 
@@
 We have applied a windowing technique to learn a GP on the data.
-We use the covariance matrix built using data from $t = 0$ till $t = \mathcal{T}$, to predict the change in stock price at time instant $t = \mathcal{T}+1$.
-Once the prediction is made, we use the input vector at $t = \mathcal{T}+1$ and the \textbf{\textit{predicted}} change in price at that instant as a new data point and compute the covariance matrix again for the window from $t = 1$ till $t = \mathcal{T}+1$.
-We use this newly computed covariance matrix to predict the change in stock price at time instant $t=\mathcal{T}+2$.
-This process is repeated till time instant $t = \mathcal{T}+T$, where $T$ denotes how far ahead in future we want to keep predicting the change in stock prices.
-Since re-computing the covariance matrix for every window can be pretty time consuming, we compute only the last column / row of the covariance matrix as it corresponds to the time instant $t = \mathcal{T}+1$.
-In addition, we drop the row and column that correspond to the oldest time instant of the previous window, from the covariance matrix.
-This keeps the size of our covariance matrix constant (therefore, keeping the amount of computation required for inverting the covariance matrix constant across every window) while allowing us to re-use the other columns of the covariance matrix that correspond to the time instants common to the previous window and the current window.
+We use the covariance matrix built using data from $t = 0$ to $t = \mathcal{T}$, to predict the change in stock price at time $t = \mathcal{T}+1$.
+Once the prediction is made, we use the input vector at $t = \mathcal{T}+1$ and the \textbf{\textit{predicted}} change in price at that instance as a new data point and compute the covariance matrix again for the window from $t = 1$ to $t = \mathcal{T}+1$.
+We use this newly computed covariance matrix to predict the change in stock price at $t=\mathcal{T}+2$.
+This process is repeated until $t = \mathcal{T}+T$, where $T$ denotes how far ahead in future we want to keep predicting the change in stock prices.
+Since re-computing the covariance matrix for every window can be time consuming, we compute only the last column / row of the covariance matrix as it corresponds to the time instance $t = \mathcal{T}+1$.
+In addition, we drop the row and column that correspond to the oldest time instance of the previous window, from the covariance matrix.
+This keeps the size of our covariance matrix constant (therefore, keeping the amount of computation required for inverting the covariance matrix constant across every window) while allowing us to re-use the other columns of the covariance matrix that correspond to the time instances common to the previous window and the current window.
 
@@
 
-We predict the change in stock price at the time instant immediately following our current time window, to be the \textbf{\textit{mean}} of the below conditional distribution:
+We predict the change in stock price at the time instance immediately following our current time window, to be the \textbf{\textit{mean}} of the below conditional distribution:
 
@@
 
-The results from our experiments indicate that GP models could learn the correlations hidden among financial data to a very high degree when compared to other traditional ML algorithms like linear regression, PCA regression and neural networks.
+The results from our experiments indicate that GP models could learn the correlations hidden among financial data to a very high degree when compared to other traditional ML algorithms like linear regression and PCA regression.
 But the biggest drawback of GP models as implemented in our application is the computation time that is required for the GP modeling.
@@
 These computational needs, restricted the amount of training data that could be used, to a maximum of only 20000 training samples, even though we had access to around 2GB of stock price data that we would have liked to use during training.
-Several authors have suggested various sparse approximation approaches towards calculating the covariance matrix in which they try to retain most of the information contained in the full dataset while reducing the size of the covariance matrix so as to facilitate a less demanding implementation of the GP model from the computational point of view.
+Several authors have suggested sparse approximation approaches for calculating the covariance matrix, which we believe is a promising direction for future work.
 
@@
 \subsubsection*{References}
-
 \small{

### 2014-05-10 paper/paper.tex
latest status

@@
 \subsection{Pre-processing}
+\label{sec:strategy}
 
@@
 \begin{align}
-	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- \frac{1}{2} u^2} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Squared Exponential}   \\
-	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Exponential}   \\
-	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{3} u \right) \exp{- \sqrt{3}  u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{3}{2}$) }   \\
-	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{5} u \right) + 5 u^2) \exp{- \sqrt{5} u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{Matern ($\nu = \frac{5}{2}$) }
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- \frac{1}{2} u^2} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Squared Exponential)}   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Ornstein–Uhlenbeck)}   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{3} u \right) \exp{- \sqrt{3}  u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Matern - $\nu = \frac{3}{2}$)}   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{5} u \right) + 5 u^2) \exp{- \sqrt{5} u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Matern  - $\nu = \frac{5}{2}$)}
 \end{align}
@@
 
+\section{Experiments}
+
+\subsection{Covariance Type}
+For this experiment, we use $\mathcal{T} = 20000$, with Strategy 3, and try to predict 500 ms into the future.
+We found that squared exponential and the Matern function with $\nu = \frac{5}{2}$ predicted nearly a straight line, with very small variations.
+The Matern functio wiht $\nu = \frac{3}{2}$ and Ornstein–Uhlenbeck functions' predictions are shaped more like the actual data, with sharp peaks and valleys.
+
+The Ornstein-Uhlenbeck model seems to fit the data the best, so we use it for all future experiments.
+
+% Zoom plot in - 50-100??
+
+\subsection{Window size}
+Next, we experimented with varying window size $\mathcal{T}$, while using Strategy 3 and predicting 500 ms into the future.
+We found that the larger windows predicted the dominant trend much better, and had fewer peaks and valleys.
+However, the peaks and valleys that the larger windows predicted were, on average, more likely to be correct.
+
+\subsection{Strategy Type}
+We tried several different strategies (described in \ref{sec:strategy}).
+\begin{table}[h]
+	\label{t:miss}
+	\centering
+	\caption{Earnings from different strategies}
+	\begin{tabular}{| l | l |}
+	\hline
+	Strategy & Earnings \\
+	\hline
+	1 & 42.88 \\
+	2 & 28.39 \\
+	3 & 3.86 \\
+	\hline
+	\end{tabular}
+\end{table}
+
+\subsection{Future time}
+In experiments up to this point, we used data that had been processed so that the y values are the change in stock price 500 ms in the future.
+Next, we re-processed the data to have y values 800 ms or 1000 ms in the future.
+We found that we got roughly the same results for each experiment, suggesting that we are equally capable of predicting on each of these time scales into the future.
+
+If we had more time, we would run additional experiments further into the future, to see how far ahead we can predict before our predictions get notably worse.
+
+\subsection{Comparisons to other models}
+The dataset that we used came with linear regression and PCA regression models, trained on a much larger version of the same dataset.
+Although the experiments are not trained and tested in the same way, we give a broad comparison to their results, to show that we performed similarly well.
+
 \section{Conclusion}
-
 The results from our experiments indicate that GP models could learn the correlations hidden among financial data to a very high degree when compared to other traditional ML algorithms like linear regression and PCA regression.

### 2014-05-10 code/graphs.m

@@ -21,14 +21,16 @@ close all;
 
 plotNum = 1;
 folder = '../data/data_20000_500_500_0/';
-files = cell(1,3);
+files = cell(1,4);
 files{1} = 'Predicted_yValues_20000_1trained_with_500vals__0.txt';
-files{2} = 'Predicted_yValues_20000_3trained_with_500vals__0.txt';
-files{3} = 'Predicted_yValues_20000_4_trained_with_500vals_rand_0.txt';
-labels = cell(1,3);
+files{2} = 'Predicted_yValues_20000_2trained_with_500vals__0.txt';
+files{3} = 'Predicted_yValues_20000_3trained_with_500vals__0.txt';
+files{4} = 'Predicted_yValues_20000_4_trained_with_500vals_rand_0.txt';
+labels = cell(1,4);
 labels{1} = '10 * Predictions with Squared Exponential';
-labels{2} = '10 * Predictions with Matern \nu = 3/2';
-labels{3} = '10 * Predictions with Matern \nu = 5/2';
+labels{2} = '10 * Predictions with Ornstein-Ulenbeck';
+labels{3} = '10 * Predictions with Matern \nu = 3/2';
+labels{4} = '10 * Predictions with Matern \nu = 5/2';
 titl = 'Different covariance functions, trained with 20000 samples';
 xlabl = 'Testing Value #';
 ylabl = 'Change in price';
@@ -63,15 +65,21 @@ if exist([p, 'processed_data.txt'], 'file')
     plot (x, y_result','b', ...
         x, 10*yPredicted(1,:),'--c', ...
         x, 10*yPredicted(2,:),'--r', ...
-        x, 10*yPredicted(3,:),'--g'); %, ...
-        % x, 10*yPredicted(3,:), 'y*');
+        x, 10*yPredicted(3,:),'--g', ...
+        x, 10*yPredicted(4,:), '--y');
         
     title(titl)
     xlabel(xlabl)
     ylabel(ylabl)
-    legend('Test values', labels{1}, labels{2}, labels{3})%, labels{4})
+    legend('Test values', labels{1}, labels{2}, labels{3}, labels{4})
 
     saveas(gcf, ['fig_', num2str(plotNum), '.png']);
+    
+    plot ( ...
[…]

### 2014-05-10 paper/paper.tex
Paper author changes

--- 2014-05-10 paper/paper.tex
@@
 \usepackage{amsmath,amssymb}
+\usepackage[T1]{fontenc}
+%\usepackage{babel}
+
 %\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
-
-\title{Applying Gaussian Process Regression to Predict Changes in Stock Price}
 
@@
 \texttt{sdewet@cs.wisc.edu} \\
-\And
+\and
 Saswati De \\
@@
 }
+
+\title{Applying Gaussian Process Regression to Predict Changes in Stock Price}
 
@@
 \begin{abstract}
+Predicting stock prices is an interesting problem for machine learning researchers.
+In this paper we use Gaussian Processes to model the timeseries representing the change in price of a single stock.
+Our implementation of the GP model performed well in comparison to other regression models like Linear Regression and PCA Regression.
 \end{abstract}
@@
 \section{Initial Data Processing}
-Our data set consists of information collected on a single stock, including instantaneous stock prices and instantaneous values of 27 parameters ($P$) known to have some affect on the price.
+Our data set (obtained from http://www.tworoads.co.in/trading-competition) consists of information collected on a single stock, including instantaneous stock prices and instantaneous values of 27 parameters ($P$) known to have some affect on the price.
 These values are taken at many different time stamps.
@@
 \subsubsection*{Acknowledgments}
-Thanks to Jerry Zhu for an excellent class, and to Rasmussen and Willams for their excellent book.
+Thanks to Prof. Zhu for an excellent class.
 
@@
 \small{
-}
+C.E. Rasmussen \& C.K.I. Williams, Gaussian Processes for Machine Learning. \\
+David Mackay, Introduction to Gaussian Processes.\\
+Long-term Stock Market Forecasting using Gaussian Processes, Nando de Freitas.
+Dataset - http://www.tworoads.co.in/trading-competition}
 

### 2014-05-10 paper/paper.tex
Final paper

@@
 \usepackage{amsmath,amssymb}
+\usepackage{graphicx}
 \usepackage[T1]{fontenc}
@@
 The majority of existing stock market forecasting techniques rely on machine learning tools and the accuracy of these tools is still a significant challenge for market traders considering the several external factors which affect the stock market.
-In this project we have tried to forecast change in price of a single stock using the Gaussian Processes (GP) model.
+In this project we have tried to forecast change in stock price using the Gaussian Processes (GP) model.
 
 \section{Initial Data Processing}
-Our data set (obtained from http://www.tworoads.co.in/trading-competition) consists of information collected on a single stock, including instantaneous stock prices and instantaneous values of 27 parameters ($P$) known to have some affect on the price.
+Our data set (obtained from [4]) consists of information collected on a single stock, including instantaneous stock prices and instantaneous values of 27 parameters ($P$) known to have some effect on the price.
 These values are taken at many different time stamps.
@@
 
-Our algorithm could be divided into 4 major parts - preprocessing the data, building the covariance matrix, tuning the hyperparameters, learning the GP model using a windowing technique and finally predicting the change in stock price for the next time instance using the learned GP model.
+Our algorithm could be divided into 5 major parts - preprocessing the data, building the covariance matrix, tuning the hyperparameters, learning the GP model using a windowing technique and finally predicting the change in stock price for the next time instance using the learned GP model.
 
@@
 We experiment with 3 different strategies to build our training set from the available dataset.
-In the first strategy we prune out all data points that were more than one standard deviation away from the sample mean $\bf\bar{y}$.
-In the second strategy we prune out all data points that were within one standard deviation from the sample mean $\bf\bar{y}$.
+In the first strategy we prune out all data points that were more than one standard deviation \textbf{away} from the sample mean $\bf\bar{y}$.
+In the second strategy we prune out all data points that were \textbf{within} one standard deviation from the sample mean $\bf\bar{y}$.
 And for the third strategy, we keep all data points.
@@
 \begin{align}
-	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- \frac{1}{2} u^2} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Squared Exponential)}   \\
-	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp{- u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Ornstein–Uhlenbeck)}   \\
-	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{3} u \right) \exp{- \sqrt{3}  u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Matern - $\nu = \frac{3}{2}$)}   \\
-	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{5} u \right) + 5 u^2) \exp{- \sqrt{5} u} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Matern  - $\nu = \frac{5}{2}$)}
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp\{- \frac{1}{2} u^2\} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Squared Exponential)}   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \exp\{- u\} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Ornstein Uhlenbeck)}   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{3} u \right) \exp\{- \sqrt{3}  u\} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Matern - $\nu = \frac{3}{2}$)}   \\
+	k(\bfx_p, \bfx_q) &= \sigma_f^2 \left( 1 + \sqrt{5} u \right) + 5 u^2) \exp\{- \sqrt{5} u\} + \sigma_n \delta_{pq} \quad \quad \quad \mbox{(Matern  - $\nu = \frac{5}{2}$)}
 \end{align}
@@
 \begin{equation}
-	\log p(\bfy | X, \bftheta) = - \frac{1}{2} \bfy^\top K_y^{-1} \bfy - \frac{1}{2} \log \| K_y \| - \frac{n}{2} \log (2 \pi)
+	\log p(\bfy | X, \bftheta) = - \frac{1}{2} \bfy^\top K_y^{-1} \bfy - \frac{1}{2} \log | K_y | - \frac{n}{2} \log (2 \pi)
 \end{equation}
@@
 
-We observed that the resultant $\bfl$ parameters had very large values in the majority of their coefficients and smaller values for a limited subset of components.
+We also observed that the resultant $\bfl$ parameters had very large values in the majority of their coefficients and smaller values for a limited subset of components.
 This means that a limited number of components have a larger correlation with the target value.
@@
 This process is repeated until $t = \mathcal{T}+T$, where $T$ denotes how far ahead in future we want to keep predicting the change in stock prices.
+
 Since re-computing the covariance matrix for every window can be time consuming, we compute only the last column / row of the covariance matrix as it corresponds to the time instance $t = \mathcal{T}+1$.
 In addition, we drop the row and column that correspond to the oldest time instance of the previous window, from the covariance matrix.
+We repeat this procedure for every new $t$ that we predict.
 This keeps the size of our covariance matrix constant (therefore, keeping the amount of computation required for inverting the covariance matrix constant across every window) while allowing us to re-use the other columns of the covariance matrix that correspond to the time instances common to the previous window and the current window.
@@
 For this experiment, we use $\mathcal{T} = 20000$, with Strategy 3, and try to predict 500 ms into the future.
-We found that squared exponential and the Matern function with $\nu = \frac{5}{2}$ predicted nearly a straight line, with very small variations.
-The Matern functio wiht $\nu = \frac{3}{2}$ and Ornstein–Uhlenbeck functions' predictions are shaped more like the actual data, with sharp peaks and valleys.
+In figure \ref{fig1}, we found that squared exponential, Ornstein Uhlenbeck and the Matern function with $\nu = \frac{5}{2}$ predicted nearly a straight line, with very small variations.
+The Matern function with $\nu = \frac{3}{2}$ had predictions shaped more like the actual data, with sharp peaks and valleys.
 
-The Ornstein-Uhlenbeck model seems to fit the data the best, so we use it for all future experiments.
+\begin{figure}[h]
+	\label{fig1}
+	\centering
+	\includegraphics[width=0.8\textwidth]{../code/fig_1.png}
+	\caption{Covariance Functions Experiment}
+\end{figure}
 
-% Zoom plot in - 50-100??
+The Matern function with $\nu = \frac{3}{2}$ looked most like the data, but the Ornstein-Uhlenbeck model had a smaller mean squared error, so we used it for all future experiments.
 
@@
 Next, we experimented with varying window size $\mathcal{T}$, while using Strategy 3 and predicting 500 ms into the future.
-We found that the larger windows predicted the dominant trend much better, and had fewer peaks and valleys.
-However, the peaks and valleys that the larger windows predicted were, on average, more likely to be correct.
+In figure \ref{fig2}, we found that the larger windows predicted the dominant trend much better, and had fewer peaks and valleys.
+
+\begin{figure}[h]
+	\label{fig2}
+	\centering
+	\includegraphics[width=0.8\textwidth]{../code/fig_2.png}
+	\caption{Window Experiments}
+\end{figure}
 
@@
 We tried several different strategies (described in \ref{sec:strategy}).
+We then applied an earnings predictor, provided with [4], to our predicted values.
+The earnings of a model measure how well it fits the dataset.
 \begin{table}[h]
-	\label{t:miss}
+	\label{t}
 	\centering
@@
 
+In table \ref{t}, we found that breaking the data into inliers and outliers helped us to predict the trend much better.
+We think this is because these models had less variation and were easier to train.
+With more training data and time, the Strategy 0 model may have done better too.
+
 \subsection{Future time}
@@
 
-If we had more time, we would run additional experiments further into the future, to see how far ahead we can predict before our predictions get notably worse.
+If we had more time, we would run additional experiments further into the future, to see how far ahead we can predict before our model's performance gets notably worse.
+We also discovered that the data does not have a high enough precision.  In many cases, the stock prices aren't given to enough decimal places, so the change in price is just 0.
+Using larger time gaps might have helped to ameliorate this problem.
 
 \subsection{Comparisons to other models}
-The dataset that we used came with linear regression and PCA regression models, trained on a much larger version of the same dataset.
-Although the experiments are not trained and tested in the same way, we give a broad comparison to their results, to show that we performed similarly well.
+The dataset that we used came with pre-trained linear regression and PCA regression models.
+In Figure \ref{fig4}, we show comparisons of earnings of the different learning algorithms.  Each group of bars is tested on the same set of data.
+
+\begin{figure}[h]
+	\label{fig4}
+	\centering
+	\includegraphics[width=0.8\textwidth]{../code/fig_4.png}
+	\caption{Comparisons Between Different Learning Algorithms}
+\end{figure}
 
@@
 Several authors have suggested sparse approximation approaches for calculating the covariance matrix, which we believe is a promising direction for future work.
+Another improvement would be to select the predicted values by sampling from the distribution, rather than using the distribution mean.
+
+\subsubsection*{Contributions}
+Saswati contributed the experiment design and the Gaussian Processes code.
+Stephanie contributed the covariance optimization code.
+Both team members ran experiments, contributed to the paper, and researched the models and methods to be used.
 
@@
 \small{
-C.E. Rasmussen \& C.K.I. Williams, Gaussian Processes for Machine Learning. \\
-David Mackay, Introduction to Gaussian Processes.\\
-Long-term Stock Market Forecasting using Gaussian Processes, Nando de Freitas.
-Dataset - http://www.tworoads.co.in/trading-competition}
+	[1] C.E. Rasmussen \& C.K.I. Williams, Gaussian Processes for Machine Learning.
+
+	[2] David Mackay, Introduction to Gaussian Processes.
+
+ [3] Long-term Stock Market Forecasting using Gaussian Processes, Nando de Freitas.
+
+ [4] Dataset - http://www.tworoads.co.in/trading-competition
+ }
 

### 2014-05-10 code/graphs.m

@@ -50,36 +50,32 @@ if exist([p, 'processed_data.txt'], 'file')
     y_training = y(1:n_train_data);
     y_result = y(n_train_data+1:n_train_data+500);
     
-    plot (1:n_train_data, y_training,'k');
-    title('Training data');
-    ylabel('Change in stock price');
-    saveas(gcf, 'fig_0.png');
-    close;
-    
+%     plot (1:n_train_data, y_training,'k');
+%     title('Training data');
+%     ylabel('Change in stock price');
+%     saveas(gcf, 'fig_0.png');
+%     close;
+%     
     next = 0;
     for j = 1 : numel(files)
         next = next + 1;
         yPredicted(next,:) = readYPred([p, files{j}]);
     end
     x = 1:500;
-    plot (x, y_result','b', ...
-        x, 10*yPredicted(1,:),'--c', ...
+    plot (x, y_result','k', ...
+        x, 10*yPredicted(4,:),'--c', ...
         x, 10*yPredicted(2,:),'--r', ...
-        x, 10*yPredicted(3,:),'--g', ...
-        x, 10*yPredicted(4,:), '--y');
+        x, 10*yPredicted(3,:),'--b', ...
+        x, 10*yPredicted(1,:), '--g');
         
     title(titl)
     xlabel(xlabl)
     ylabel(ylabl)
+    xlim([200 400])
+    ylim([-1 1.5])
     legend('Test values', labels{1}, labels{2}, labels{3}, labels{4})
 
     saveas(gcf, ['fig_', num2str(plotNum), '.png']);
-    
-    plot ( ...
-        x, log((y_result' - yPredicted(1,:)).^2),'--c', ...
-        x, log((y_result' - yPredicted(2,:)).^2),'--r', ...
-        x, log((y_result' - yPredicted(3,:)).^2),'--g', ...
-        x, log((y_result' - yPredicted(3,:)).^2),'--y');
 end
 
 
@@ -103,7 +99,7 @@ xlabl = 'Testing Value #';
 ylabl = 'Change in price';
 
 next = 0;
-for j = 1 : 2%numel(folders)
+for j = 1 : numel(folders)
     next = next + 1;
     p = ['../data/', folders{j}, '/']; % path
     [y, ~] = readProcessedData(p);
[…]
